### 2009-11-24 trunk/Paper/paper.tex
Added Paper

\documentclass[letterpaper]{article}
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PDFMARK for TeX and GhostScript
% Uncomment and complete the following for metadata if
% your paper is typeset using TeX and GhostScript (e.g
% if you use .ps or .eps files in your paper):
% \special{! /pdfmark where
% {pop} {userdict /pdfmark /cleartomark load put} ifelse
% [ /Author (John Doe, Jane Doe)
% /Title (Paper Title)
% /Keywords (AAAI, artificial intelligence)
% /DOCINFO pdfmark}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PDFINFO for PDFTeX
% Uncomment and complete the following for metadata if
% your paper is typeset using PDFTeX
\pdfinfo{
  /Title (Solving an Exam Scheduling Problem Using a Genetic Algorithm)
  /Subject (Input The Proceedings Title Here)
  /Author (Dave, Kordalewski;
           Caigu, Liu;
           Kevin, Salvesen;)
} 
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Uncomment only if you need to use section numbers
% and change the 0 to a 1 or 2
% \setcounter{secnumdepth}{0}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Solving an Exam Scheduling Problem Using a Genetic Algorithm}
\author{Dave Kordalewski \\ davekordalewski@gmail.com
   \And Caigu Liu \\liucaigu@gmail.com
   \And Kevin Salvesen \\ kevin.salvesen@gmail.com}

\begin{document}
\maketitle

\section{The Problem}
Description of our problem

\section{Approach and Implementation}
Tools, algorithms, APIs, etc. that we used.

\section{Evaluation}
How we determined the success of our project. Did we solve it? Effectively? Labelled images, graphs and/or tables can be useful.

% \bibliography{Bibliography-File}
% \bibliographystyle{aaai}
\end{document}

### 2009-11-24 Paper/paper.tex

--- 2009-11-24 trunk/Paper/paper.tex
@@
 
+\nocopyright
+
 \begin{document}
 \maketitle
+
+\begin{abstract}
+  This concise, one-paragraph summary should describe the general thesis
+  and conclusion of your paper. A reader should be able to learn the purpose
+  of the paper and the reason for its importance from the abstract.
+\end{abstract}
 
@@
 \section{Evaluation}
-How we determined the success of our project. Did we solve it? Effectively? Labelled images, graphs and/or tables can be useful.
+  How we determined the success of our project. Did we solve it? Effectively?
+  Labelled images, graphs and/or tables can be useful.
+
+\section{Conclusion}
+  What have you learned because of your project? What can we learn from your
+  project? Did it give you ideas as to things you might like to try next?
 

### 2009-11-24 trunk/GA_TestBench/src/ScheduleProblem/Schedule.java
added legends for the 2d graph

@@ -364,8 +364,6 @@ public class Schedule extends ga_testbench.Individual implements Cloneable {
         final Schedule otherSched = (Schedule) other;
 
         Schedule ret = null;
-
-
         try {
             ret = this.clone();
         } catch (CloneNotSupportedException ex) {
@@ -389,12 +387,7 @@ public class Schedule extends ga_testbench.Individual implements Cloneable {
                 ret.times[i] = otherSched.times[i];
             }
         }
-
-
-
         return ret;
-
-
     }
 
     /**

### 2009-11-24 trunk/Paper/paper.tex

@@
 
+\nocopyright
+
 \begin{document}
@@
 
+\begin{abstract}
+  This concise, one-paragraph summary should describe the general thesis
+  and conclusion of your paper. A reader should be able to learn the purpose
+  of the paper and the reason for its importance from the abstract.
+\end{abstract}
+
 \section{The Problem}
-Description of our problem
+  Consider a university, with all of its students, courses and classrooms. 
+  How does one try to build the best possible schedule for an exam session? 
+  A good schedule is a schedule where:
+  \begin{itemize}
+  \item  There are no conflicts, such as two different exams in the same room
+         at the same, or a student having two different exams at the same time.
+  \item  Students don't have too many exams in a very short time.
+  \item  It is bad for a room to have two exams directly one after another,
+         as it doesn't give time for teachers and TAs to set up the room.
+  \end{itemize}
+
+ Considering a huge university like University of Toronto, with its 70000 students
+ and numerous courses, making a good schedule by hand would be very tedious.
 

### 2009-11-25 GA_TestBench/src/ScheduleProblem/Main.java
Removed the GAMeanStats from getAllStats

@@ -21,7 +21,6 @@ public class Main {
     private static String randomSearchString = "";
     private static String mutateSearchString = "";
     private static String GABestStats = "";
-    private static String GAMeanStats = "";
     private static String randomSearchStats = "";
     private static String hillClimbingStats = "";
     private static String mutateSearchStats = "";
@@ -37,15 +36,15 @@ public class Main {
 //        getGAParameterStats(51, 150, 1);
 //        getGASizeStats();
 
-        getAllStats(60, 152, 1, 5, 54, 0, HUGEINSTANCE);
+//        getAllStats(60, 152, 1, 5, 54, 0, HUGEINSTANCE);
 
-//        int maxpop = 40;
-//        int maxgen = 100;
-//        double copies = 2;
-//        double mutations = 20;
-//        double crossovers = 40;
-//        double randoms = 25;
-//        getAllStats(maxpop, maxgen, copies, mutations, crossovers, randoms, SMALLINSTANCEFILE);
+        int maxpop = 40;
+        int maxgen = 100;
+        double copies = 2;
+        double mutations = 20;
+        double crossovers = 40;
+        double randoms = 25;
+        getAllStats(maxpop, maxgen, copies, mutations, crossovers, randoms, SMALLINSTANCEFILE);
     }
 
     /**
@@ -223,7 +222,6 @@ public class Main {
         int evals = worker.numEvaluations();
 
         GABestStats = worker.getBestStatsString();
-        GAMeanStats = worker.getMeanStatsString();
         return evals;
     }
 
@@ -255,12 +253,13 @@ public class Main {
             out.write("5\n");
 
             out.write(GABestStats + "\n");
-            out.write(GAMeanStats + "\n");
             out.write(randomSearchStats + "\n");
             out.write(hillClimbingStats + "\n");
             out.write(mutateSearchStats + "\n");
 
             out.close();
+
+            System.out.println("Output written in " + outFileName);
         } catch (IOException ex) {
             System.err.println("Error opening file " + outFileName + " for write");
         }

### 2009-11-26 GA_TestBench/src/ScheduleProblem/Main.java
ImplementedgetAllAverageStats() and made a few various changes around the code to make all the search string outputs more standardized

@@ -3,8 +3,11 @@ package ScheduleProblem;
 import ga_testbench.GAParams;
 import ga_testbench.GASolver;
 import java.io.BufferedWriter;
+import java.io.File;
+import java.io.FileNotFoundException;
 import java.io.FileWriter;
 import java.io.IOException;
+import java.util.Scanner;
 
 public class Main {
 
@@ -38,13 +41,13 @@ public class Main {
 
 //        getAllStats(60, 152, 1, 5, 54, 0, HUGEINSTANCE);
 
-        int maxpop = 40;
-        int maxgen = 100;
-        double copies = 2;
-        double mutations = 20;
-        double crossovers = 40;
-        double randoms = 25;
-        getAllStats(maxpop, maxgen, copies, mutations, crossovers, randoms, SMALLINSTANCEFILE);
+        int maxpop = 28;
+        int maxgen = 55;
+        double copies = 3;
+        double mutations = 4;
+        double crossovers = 100;
+        double randoms = 5;
+        getAllAverageStats(maxpop, maxgen, copies, mutations, crossovers, randoms, MEDIUMINSTANCEFILE, 10);
     }
 
     /**
@@ -186,6 +189,179 @@ public class Main {
         }
     }
 
+    public static void getAllAverageStats(int popsize, int numgens, double numCopies,
+            double numMutations, double numCrossovers, double numRandoms, String instanceFile, int numIterations) {
+
+        GAParams params = new GAParams(instanceFile, popsize, numgens, numCopies, numMutations, numCrossovers, numRandoms, true);
+
+        // We don't mind having more points for the non-GA methods
+        int samplePeriod = params.numEvalsPerGeneration();
+        
+        System.out.println(samplePeriod);
+        int maxEvals = 0;
+        for (int i = 1; i <= numIterations; i++) {
+            System.out.println("--- Running iteration number " + i + " ---");
+            // Run the 4 different searches, using the number of evaluations made by
+            // the GA search as the max number of evaluations the other algos are allowed to use
+            System.out.println("We expect " + (params.numEvalsPerGeneration() * params.getMaxGenerations()) + " evaluations to be performed.");
+            maxEvals = getGAStats(params);
+            System.out.println("GA Done! (" + maxEvals + " evals performed)");
+            getRandomSearchStats(maxEvals, samplePeriod);
+            System.out.println("Random Done!");
+            getHillClimbingStats(maxEvals, samplePeriod);
+            System.out.println("HillClimbing Done!");
[…]

### 2009-11-26 GA_TestBench/src/ScheduleProblem/Schedule.java

@@ -176,7 +176,7 @@ public class Schedule extends ga_testbench.Individual implements Cloneable {
      */
     public static Individual random() {
         verifyInitialized();
-
+        
         Timing[] times = new Timing[numCourses];
         int[] timingRooms = new int[numCourses];
         for (int i = 0; i < numCourses; i++) {
@@ -509,7 +509,10 @@ public class Schedule extends ga_testbench.Individual implements Cloneable {
      * @return
      */
     public static String getHillclimbSearchString() {
-        return hillclimbString + "END-SERIES";
+        String retString = hillclimbString + "END-SERIES";
+        hillclimbString = "";
+        hillEvals = 0;
+        return retString;
     }
 
     /**
@@ -531,7 +534,9 @@ public class Schedule extends ga_testbench.Individual implements Cloneable {
         }
         double fitness = fitness();
         hillEvals++;
-        addHillDataPoint(fitness);
+        if (hillEvals == 1)
+            addHillDataPoint(fitness);
+        
         try {
             for (int i = 0; i < numCourses; i++) {
                 for (int j = 0; j < numRooms; j++) {
@@ -549,7 +554,7 @@ public class Schedule extends ga_testbench.Individual implements Cloneable {
                             addHillDataPoint(fitness);
                         }
                         if (hillEvals >= maxEvals) {
-                            addHillDataPoint(fitness);
+                            //addHillDataPoint(fitness);
                             return best;
                         }
                     }
@@ -568,7 +573,7 @@ public class Schedule extends ga_testbench.Individual implements Cloneable {
                             addHillDataPoint(fitness);
                         }
                         if (hillEvals >= maxEvals) {
-                            addHillDataPoint(fitness);
+                            //addHillDataPoint(fitness);
                             return best;
                         }
                     }
@@ -587,7 +592,7 @@ public class Schedule extends ga_testbench.Individual implements Cloneable {
                             addHillDataPoint(fitness);
                         }
                         if (hillEvals >= maxEvals) {
-                            addHillDataPoint(fitness);
+                            //addHillDataPoint(fitness);
                             return best;
                         }
                     }
@@ -600,7 +605,7 @@ public class Schedule extends ga_testbench.Individual implements Cloneable {
 //        System.out.println("Hillclimbing made " + evals + " fitness evaluations...");
 
         if (best == this) {
-            addHillDataPoint(fitness);
+            //addHillDataPoint(fitness);
             return best;
         } else {
             return best.hillClimb(maxEvals, samplePeriod);

### 2009-11-26 GA_TestBench/src/ScheduleProblem/Main.java
just some tiny changes to kev's new method

@@ -39,15 +39,15 @@ public class Main {
 //        getGAParameterStats(51, 150, 1);
 //        getGASizeStats();
 
-//        getAllStats(60, 152, 1, 5, 54, 0, HUGEINSTANCE);
+//        getAllStats(50, 600, 1, 9, 110, 0, HUGEINSTANCE);
 
-        int maxpop = 28;
-        int maxgen = 55;
-        double copies = 3;
-        double mutations = 4;
-        double crossovers = 100;
-        double randoms = 5;
-        getAllAverageStats(maxpop, maxgen, copies, mutations, crossovers, randoms, MEDIUMINSTANCEFILE, 10);
+        int maxpop = 70;
+        int maxgen = 400;
+        double copies = 1;
+        double mutations = 6;
+        double crossovers = 63;
+        double randoms = 0;
+        getAllAverageStats(maxpop, maxgen, copies, mutations, crossovers, randoms, SMALLINSTANCEFILE, 10);
     }
 
     /**
@@ -196,14 +196,14 @@ public class Main {
 
         // We don't mind having more points for the non-GA methods
         int samplePeriod = params.numEvalsPerGeneration();
-        
-        System.out.println(samplePeriod);
+
+        System.out.println("The sample period is " + samplePeriod);
         int maxEvals = 0;
         for (int i = 1; i <= numIterations; i++) {
-            System.out.println("--- Running iteration number " + i + " ---");
+            System.out.println("\n--- Running iteration number " + i + " of " + numIterations + " ---");
             // Run the 4 different searches, using the number of evaluations made by
             // the GA search as the max number of evaluations the other algos are allowed to use
-            System.out.println("We expect " + (params.numEvalsPerGeneration() * params.getMaxGenerations()) + " evaluations to be performed.");
+            System.out.println("We expect " + (params.numEvalsPerGeneration() * params.getMaxGenerations() + params.getPopSize()) + " evaluations to be performed.");
             maxEvals = getGAStats(params);
             System.out.println("GA Done! (" + maxEvals + " evals performed)");
             getRandomSearchStats(maxEvals, samplePeriod);
@@ -235,11 +235,11 @@ public class Main {
         String HillTitle = "";
         String MutateTitle = "";
 
-        int[] fitnessIntervals = new int[maxEvals/samplePeriod+1];
-        double[] GAStatsSum = new double[maxEvals/samplePeriod+1];
[…]

### 2009-11-26 GA_TestBench/src/ScheduleProblem/Main.java
Fixed getAllAverageStats() to use a random number in output files, so that we can run it several times at the same time

@@ -8,6 +8,7 @@ import java.io.FileNotFoundException;
 import java.io.FileWriter;
 import java.io.IOException;
 import java.util.Scanner;
+import java.util.Random;
 
 public class Main {
 
@@ -40,14 +41,14 @@ public class Main {
 //        getGASizeStats();
 
 //        getAllStats(50, 600, 1, 9, 110, 0, HUGEINSTANCE);
-
-        int maxpop = 70;
-        int maxgen = 400;
+        
+        int maxpop = 30;
+        int maxgen = 40;
         double copies = 1;
         double mutations = 6;
         double crossovers = 63;
         double randoms = 0;
-        getAllAverageStats(maxpop, maxgen, copies, mutations, crossovers, randoms, SMALLINSTANCEFILE, 10);
+        getAllAverageStats(maxpop, maxgen, copies, mutations, crossovers, randoms, SMALLINSTANCEFILE, 5);
     }
 
     /**
@@ -192,6 +193,8 @@ public class Main {
     public static void getAllAverageStats(int popsize, int numgens, double numCopies,
             double numMutations, double numCrossovers, double numRandoms, String instanceFile, int numIterations) {
 
+        Random rand = new Random();
+        int randomInstanceNumber = rand.nextInt();
         GAParams params = new GAParams(instanceFile, popsize, numgens, numCopies, numMutations, numCrossovers, numRandoms, true);
 
         // We don't mind having more points for the non-GA methods
@@ -214,21 +217,21 @@ public class Main {
             System.out.println("MutateSearch Done!");
 
             // Generate temporary files holding data of each iteration
-            String outfile = "Data Files\\tempAverageStatsOutput\\" + i + ".txt";
+            String outfile = "Data Files\\tempAverageStatsOutput\\" + randomInstanceNumber + "_" + i + ".txt";
             printStats(outfile, maxEvals);
         }
 
         // We generate uniform filenames so that it's easy to know what data is in a file
-        String outfile = "Data Files\\FullAverageStats over " + numIterations + " iterations with " + maxEvals + " evals " +
+        String outfile = "Data Files\\FullAverageStats(" + randomInstanceNumber + ") over " + numIterations + " iterations with " + maxEvals + " evals " +
                 params.getPopSize() + " pop " + params.getMaxGenerations() + " gens " +
                 params.getNumCopies() + " copies " + params.getNumMutations() + " mutations " +
[…]

### 2009-11-26 GA_TestBench/src/ScheduleProblem/Main.java
Updated get all stats to force hillclimbing to continue until attaining the max number of evals, some loss of processing, but results in more consistent output files. There is still a small glitch in HillClimbing: it seems to skip one value somewhere, meaning that the last point in the final average output file will have fitness 0 for hillcliming. Is not important, as it will have attained its ceiling way before that. Might try to fix that later...

@@ -42,13 +42,13 @@ public class Main {
 
 //        getAllStats(50, 600, 1, 9, 110, 0, HUGEINSTANCE);
         
-        int maxpop = 30;
-        int maxgen = 40;
-        double copies = 1;
-        double mutations = 6;
-        double crossovers = 63;
+        int maxpop = 125;
+        int maxgen = 200;
+        double copies = 0;
+        double mutations = 0;
+        double crossovers = 100;
         double randoms = 0;
-        getAllAverageStats(maxpop, maxgen, copies, mutations, crossovers, randoms, SMALLINSTANCEFILE, 5);
+        getAllAverageStats(maxpop, maxgen, copies, mutations, crossovers, randoms, SMALLINSTANCEFILE, 3);
     }
 
     /**
@@ -249,6 +249,7 @@ public class Main {
         for (int i = 1; i <= numIterations; i++) {
             File infile = new File("Data Files\\tempAverageStatsOutput\\" + instanceNumber + "_" + i + ".txt");
             Scanner scanner = null;
+            String temp = null;
             try {
                 scanner = new Scanner(infile);
             } catch (FileNotFoundException ex) {
@@ -265,50 +266,54 @@ public class Main {
                 }
             }
 
-            GATitle = scanner.nextLine() + "\n";
             // Get GA data and the intervals if this is first pass through
-            for (int j = 1; j <= maxEvals / samplePeriod; j++) {
-                String temp = scanner.nextLine();
+            GATitle = scanner.nextLine() + "\n";
+            temp = scanner.nextLine();
+            int j = 1;
+            while (!temp.contentEquals("END-SERIES")) {
                 String[] vals = temp.split(", ");
-
                 if (i == 1) {
                     fitnessIntervals[j] = Integer.parseInt(vals[0]);
                 }
                 GAStatsSum[j] += Double.parseDouble(vals[1]);
+                temp = scanner.nextLine();
+                j++;
             }
 
-            scanner.nextLine(); // Skip "END-SERIES"
-
-
-            RandomTitle = scanner.nextLine() + "\n";
             // Get Random data
-            for (int j = 0; j <= maxEvals / samplePeriod; j++) {
-                String temp = scanner.nextLine();
+            RandomTitle = scanner.nextLine() + "\n";
+            temp = scanner.nextLine();
+            j = 0;
+            while (!temp.contentEquals("END-SERIES")) {
                 String[] vals = temp.split(", ");
[…]

### 2009-11-26 trunk/GA_TestBench/src/ScheduleProblem/Schedule.java
Updated get all stats to force hillclimbing to continue until attaining the max number of evals, some loss of processing, but results in more consistent output files. There is still a small glitch in HillClimbing: it seems to skip one value somewhere, meaning that the last point in the final average output file will have fitness 0 for hillcliming. Is not important, as it will have attained its ceiling way before that. Might try to fix that later...

@@ -604,7 +604,8 @@ public class Schedule extends ga_testbench.Individual implements Cloneable {
 
 //        System.out.println("Hillclimbing made " + evals + " fitness evaluations...");
 
-        if (best == this) {
+        //if (best == this) {
+        if (hillEvals >= maxEvals) {
             //addHillDataPoint(fitness);
             return best;
         } else {

### 2009-11-27 trunk/Paper/paper.tex
Updated paper with what Dave wrote on the google doc so we get a rough idea of how much we have and what the final paper looks like

--- 2009-11-24 trunk/Paper/paper.tex
@@
 
+\begin{abstract}
+  This concise, one-paragraph summary should describe the general thesis
+  and conclusion of your paper. A reader should be able to learn the purpose
+  of the paper and the reason for its importance from the abstract.
+\end{abstract}
+
 \section{The Problem}
-Description of our problem
+  The problem we investigate is a sort of scheduling problem. 
+  We are attempting to find the most satisfactory choice of when
+  and where to hold exams, given a background of student course loads.
+
+  An instance of this scheduling problem consists of a number of days
+  on which exams can be scheduled ($d$), the number of time slots in which
+  an exam can be scheduled on any day ($t$), a set of rooms ($R$), a set of
+  courses ($C$), and a set of students ($S$), each of whom ($s$) has a
+  particular course load, some subset of the courses ($L_s$).
+
+  A schedule, then, is a mapping from courses to rooms at times, 
+  which we can express like this:
+
+  \[ C \rightarrow \left\{(r,\ a,\ b)\ |\ r\in R,\ a\in \{1..d\},\ b\in \{1..t\}\right\} \]
+
+  The total number of different possible schedules in any problem instance is
+  $|C|^{|R|dt}$ which is typically far too large for brute force search.
+  For example, one instance that we examine (scheduling the fall semester
+  exams at the University of Toronto in 2009) involves 603 courses, 7 days,
+  8 times, 43 rooms, and 21945 students. This instance admits approximately 
+  $10^{6695}$ different possible schedules.
+
+  Typically, with relatively loose constraints on the number of rooms and 
+  particular schedules of students, it is not difficult to find some consistent
+  schedule; that is, one where no student is asked to write 2 exams
+  simultaneously and no room has 2 exams occur in it simultaneously.
+
+  Rather than worry about hard constraints, we prefer a framework of soft 
+  constraints, where we try to find a schedule that makes the students and 
+  invigilators happiest overall, allowing the possibility that some student 
+  or room is left with an impossible exam schedule, which can, in the real 
+  world, be dealt with on an individual basis.
+
+  The timetable that a student $s \in S$ has under some particular schedule 
+  $K$ may be represented in this way:
+
+  \[ TT_s(K) = \left\{(a, b)\ |\ \exists r \in R, \exists c \in L_s, K(c)=(r, a, b) \right\} \]
+
+  which may, in general, be a multiset.
+
+  Similarly, the timetable for a room $r \in R$ is
+
+  \[ TT_r(K)=\left\{(a, b)\ |\ \exists c\in C,\ K(c)=(r, a, b)\right\} \]
+
+  We define two quality functions, mapping schedules to real numbers in [0,1],
+  one for students and one for rooms. These are meant to capture how much they
+  "like" the schedule under consideration. (For instance, a student will rate
+  poorly any schedule where she has 2 consecutive exams, and very poorly any 
+  schedule which expects her to take two exams at the same time.)
+
+  \[ Q_s(K)=f\left(TT_s(K)\right):\textit{ Schedules}\rightarrow [0,\ 1] \]
+  \[ Q_r(K)=g\left(TT_r(K)\right):\textit{ Schedules}\rightarrow [0,\ 1] \]
+
+  We will consider later how to define these functions in a reasonable way.
+
+  The quality of a schedule, then, is given by
+
+  \[ Q(K)= \frac{\left(w_s\left(\sum_{s \in S}{\frac{Q_s(K)}{|S|}}\right) +
+     w_r\left(\sum_{r\in R}{\frac{Q_r(K)}{|R|}} \right)\right)}{(w_s+w_r)} \]
+
+  where $w_s$ and $w_r$ are the weightings we can use to indicate that we 
+  care somewhat more about students preferences than rooms, giving a quality
+  (or fitness, in the terminology of genetic algorithms) in the range [0,1] for
+  any schedule.
+
+  Evaluating this function Q can be computationally expensive when $|S|$ and $|R|$
+  are large. Our work examines how to find a schedule with relatively high fitness
+  considering that we will want to do this with as few evaluations of Q as possible.
+
+  We examine, first and foremost, a genetic algorithm approach to solving this problem,
+  but also touch on a few other methods.
 

### 2009-11-27 GA_TestBench/src/ScheduleInstanceGenerator/Main.java
New method for generating instance files done, String outFilePrepend field added to tester functions, to allow us to attach some info to the automatically generated output data file names, making it easier to track the importance of the output, so long as it is set correctly before data runs

@@ -15,7 +15,7 @@ import java.util.TreeSet;
  */
 public class Main {
 
-    public static final String TESTINSTANCE = "Instance Input Files\\tickle.txt";
+    public static final String TESTINSTANCE = "Instance Input Files\\testing instance.txt";
 //    private static final String FILENAME = "Instance Input Files\\smallish instance.txt";
     public static final String FILENAME = TESTINSTANCE;
     // If this is set to true, we will generate an instance the size of the sun
@@ -24,12 +24,21 @@ public class Main {
     private static final String[] trueCourseNames = {"ABS353H1F", "ACT230H1F", "ACT240H1F", "ACT348H1F", "ACT349H1F", "ACT451H1F", "ACT460H1F", "ACT460H1F", "ANA300Y1Y", "ANT204H1F", "ANT319Y1Y", "ANT345H1F", "ANT364Y1Y", "APM236H1F", "APM346H1F", "APM421H1F", "ARC131H1F", "ARC131H1F", "ARC335H1F", "AST101H1F", "AST101H1F", "AST101H1F", "AST101H1F", "AST101H1F", "AST101H1F", "AST101H1F", "AST101H1F", "AST210H1F", "BCH210H1F", "BCH210H1F", "BCH210H1F", "BCH210H1F", "BCH210H1F", "BCH210H1F", "BCH242Y1Y", "BCH335H1F", "BCH370H1F", "BCH371H1Y", "BCH422H1F", "BCH425H1F", "BCH441H1F", "BCH446H1F", "BCH471Y1Y", "BIO150Y1Y", "BIO240H1F", "BIO240H1F", "BIO240H1F", "BIO240H1F", "BIO240H1F", "BIO240H1F", "BIO240H1F", "BIO251Y1Y", "BIO255Y1Y", "BIO270H1F", "BIO270H1F", "CHM138H1F", "CHM139H1F", "CHM151Y1Y", "CHM217H1F", "CHM220H1F", "CHM220H1F", "CHM220H1F", "CHM247H1F", "CHM326H1F", "CHM338H1F", "CHM342H1F", "CHM347H1F", "CHM348H1F", "CHM410H1F", "CHM417H1F", "CHM427H1F", "CHM432H1F", "CHM434H1F", "CHM440H1F", "CHM441H1F", "CHM447H1F", "CLA160H1F", "CLA204H1F", "CLA219H1F", "CLA230H1F", "CLA230H1
[…]

### 2009-11-27 trunk/GA_TestBench/src/ScheduleProblem/Main.java
New method for generating instance files done, String outFilePrepend field added to tester functions, to allow us to attach some info to the automatically generated output data file names, making it easier to track the importance of the output, so long as it is set correctly before data runs

@@ -40,15 +40,15 @@ public class Main {
 //        getGAParameterStats(51, 150, 1);
 //        getGASizeStats();
 
-//        getAllStats(50, 600, 1, 9, 110, 0, HUGEINSTANCE);
-        
-        int maxpop = 125;
-        int maxgen = 200;
-        double copies = 0;
-        double mutations = 0;
-        double crossovers = 100;
-        double randoms = 0;
-        getAllAverageStats(maxpop, maxgen, copies, mutations, crossovers, randoms, SMALLINSTANCEFILE, 3);
+        getAllStats(60, 400, 1, 4, 55, 0, MEDIUMINSTANCEFILE, "MEDIUM");
+
+//        int maxpop = 40;
+//        int maxgen = 400;
+//        double copies = 1;
+//        double mutations = 3;
+//        double crossovers = 36;
+//        double randoms = 0;
+//        getAllAverageStats(maxpop, maxgen, copies, mutations, crossovers, randoms, SMALLINSTANCEFILE, 3, "TEST");
     }
 
     /**
@@ -191,11 +191,12 @@ public class Main {
     }
 
     public static void getAllAverageStats(int popsize, int numgens, double numCopies,
-            double numMutations, double numCrossovers, double numRandoms, String instanceFile, int numIterations) {
+            double numMutations, double numCrossovers, double numRandoms, String instanceFile, int numIterations, String outFilePrepend) {
 
         Random rand = new Random();
         int randomInstanceNumber = rand.nextInt();
         GAParams params = new GAParams(instanceFile, popsize, numgens, numCopies, numMutations, numCrossovers, numRandoms, true);
+
 
         // We don't mind having more points for the non-GA methods
         int samplePeriod = params.numEvalsPerGeneration();
@@ -222,16 +223,16 @@ public class Main {
         }
 
         // We generate uniform filenames so that it's easy to know what data is in a file
-        String outfile = "Data Files\\FullAverageStats(" + randomInstanceNumber + ") over " + numIterations + " iterations with " + maxEvals + " evals " +
+        String outfile = "Data Files\\" + outFilePrepend + " " + "FullAverageStats(" + randomInstanceNumber + ") over " + numIterations + " iterations with " + maxEvals + " evals " +
                 params.getPopSize() + " pop " + params.getMaxGenerations() + " gens " +
                 params.getNumCopies() + " copies " + params.getNumMutations() + " mutations " +
                 params.getNumCrossovers() + " crossovers " + params.getNumRandoms() + " randoms" + ".txt";
 
[…]

### 2009-11-29 trunk/Paper/paper.tex
Updated paper, added bibliography

@@
 \usepackage{courier}
+\usepackage{url}
 % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \section{Approach and Implementation}
-  Tools, algorithms, APIs, etc. that we used.
+  \subsection{Differences From Natural Exam Scheduling Problems}
+    The problem we have constructed here is a compromise between the most
+    mathematically pure problem with the same sort of properties (which might ignore the
+    rooms and related constraints completely) and the problem that must be solved in the
+    real world when scheduling thousands of students, which has some important differences
+    from this problem.
+
+  \subsection{Tools used}
+    The program is implemented in Java, to generate possible schedules and to calculate
+    their fitness. A Mathematica program is also written to plot graphs, given the data
+    generated by the Java program. Additionally, we use Google code 
+    subversion\footnote{\url{http://code.google.com/p/csc384-genetic-algorithms-project/}}
+    and Google doc to share the work.
+
+  \subsection{Fitness Function}
+    In order to assign each schedule a fitness, we need to define the functions 
+    $q_{student}$ and $q_{room}$. They need to capture the idea of determining how
+    much each student or room likes their given timetable. We do this by calculating 
+    the penalty incurred by a room, which is a measure of it having properties that 
+    are unpleasant to the student or room, and then using the formula
+
+    \[ q(TT)=\frac{1}{1+penalty_{TT}} \]
+
+    This function will be in the desired range or [0,1].
+
+    For students, we accrue a penalty of 50 for every time that is repeated in a 
+    timetable, 5 when there are exams in 2 consecutive times, 3 when 2 exams happen 
+    on the same day, and 0.5 when the student has exams on consecutive days. This 
+    accords reasonably well with real students' hopes for their exam timetables.
+
+    Rooms have the same penalty of 50 when it is used by two exams at the same time, 
+    5 if it hosts exams in consecutive time slots (reasoning that the invigilators 
+    need time to let the students out and in and prepare for the next exam) and a 
+    penalty of 0.5 when a room is empty for an entire day between uses (reasoning 
+    that the room requires some amount of modifications to be used for exams, and 
+    it is difficult to use it for other purposes if it will still be used for exams 
+    soon.)
+
+  \subsection{Genetic Algorithm Search}
+    Although there are many readily available open source genetic algorithm software
+    packages available (see JGAP \cite{website:JGAP}, Jenes\cite{website:jenes}) we 
+    decided to write a simple, yet general, genetic algorithm package, in Java, that
+    can interface with data types designed by others, so long as they implement a few
+    necessary methods. The package we wrote is more than adequate for investigating
+    the scheduling problem under consideration here.
+
+    \subsubsection{General Description}
+      [\ldots]
+    \subsubsection{Mutation Operator}
+      [\ldots]
+    \subsubsection{Crossover Operator}
+      [\ldots]
+  \subsection{Other Search Algorithms}
+    \subsubsection{Random Search}
+      The random search method simply generates random schedules, evaluates their fitness,
+      and remembers the one that it has seen that has highest fitness. This method is meant
+      as a baseline against which to compare the other methods. If a search method doesn't
+      perform significantly better than random search with the same number of fitness
+      evaluations, it is useless.
+
+    \subsubsection{Hillclimbing Search}
+      The hillclimbing search that we have implemented examines every schedule in the
+      neighbourhood of a given starting schedule, and then continues from the one with
+      highest fitness, terminating when it has performed as many evaluations as it is allowed
+      to, or it has found a schedule whose neighbour are all equal or inferior to it.
+
+      The neighbouring schedules are those where only one course meets in either a different
+      room, on a different day, or at a different time. This leads to a total of
+      $ |C|(d+t+|R|-3) $ schedules in the neighbourhood of any given schedule.
+
+      It can be seen that this method is completely deterministic given a starting schedule,
+      and will always return a local maximum
+
+    \subsubsection{Mutate Search}
+      Mutate Search starts by generating a random schedule. It modifies this schedule (by
+      using the same mutate operator made for the genetic algorithm) by 1\% and rolls back if
+      the mutation didn't increase the schedule's fitness. This operation is repeated until it has
+      performed as many evaluations as it is allowed to.
+      This algorithm is a variation of simulated annealing, where instead of accepting a change
+      through a probabilistic function, it always accepts a change that leads to a higher fitness.
 
 \section{Evaluation}
-  How we determined the success of our project. Did we solve it? Effectively?
-  Labelled images, graphs and/or tables can be useful.
+  \subsection{Success}
+    [\ldots]
+    Seeing as our fitness function gives a very high penalty for schedules that have direct
+    conflicts (either a student with two exams at the same moment, or two different exams
+    in a same room at the same moment), if such a schedule exists given the input data, our
+    algorithms will very quickly tend to reject all such schedules.
+  \subsection{The Genetic Algorithm's Parameters}
+    [\ldots]
+  \subsection{Comparison Between the Different Searches}
+    [\ldots]
 
 \section{Conclusion}
-  What have you learned because of your project? What can we learn from your
-  project? Did it give you ideas as to things you might like to try next?
+  We come up with a conclusion that in this particular problem, genetic algorithm
+  is better than random search, and it is better than hill-climbing if the input 
+  is a large data set. But it may not be the best solution for this problem, since
+  the MutateSearch performs way better.
 
-% \bibliography{Bibliography-File}
-% \bibliographystyle{aaai}
+  %\cite{website:jgap} 
+  %\cite{website:jenes} 
+  %\cite{website:coderepos} 
+\bibliographystyle{aaai}
+\bibliography{paperBib}
 \end{document}

### 2009-11-30 trunk/Paper/paper.tex
Updated paper

@@
 \begin{abstract}
-  This concise, one-paragraph summary should describe the general thesis
-  and conclusion of your paper. A reader should be able to learn the purpose
-  of the paper and the reason for its importance from the abstract.
+  We examine the exam scheduling problem with soft constraints for problems of varying
+  sizes using a genetic algorithm, random search, hillclimbing, and a variant of simulated
+  annealing. This is done in the context of maximizing the fitness of a schedule, where the
+  fitness is relatively computationally expensive to compute for any schedule. The genetic
+  algorithm is discussed in detail. We consider selecting parameters for the GA that
+  maximize the quality of the schedule found after a constant number of evaluations. The
+  GA, with a naive crossover operator, performs much better than the random search, but
+  simulated annealing is far superior in this search space.
 \end{abstract}
@@
 
-  \[ Q_s(K)=f\left(TT_s(K)\right):\textit{ Schedules}\rightarrow [0,\ 1] \]
-  \[ Q_r(K)=g\left(TT_r(K)\right):\textit{ Schedules}\rightarrow [0,\ 1] \]
+  \[ Q_s(K)=q_{student}\left(TT_s(K)\right):\textit{ Schedules}\rightarrow [0,\ 1] \]
+  \[ Q_r(K)=q_{room}\left(TT_r(K)\right):\textit{ Schedules}\rightarrow [0,\ 1] \]
 
@@
 
+    The details of the function we used are essentially arbitrary.
+
   \subsection{Genetic Algorithm Search}
@@
     \subsubsection{General Description}
-      [\ldots]
+      The genetic algorithm maintains a population of schedules and constructs subsequent
+      generations by adding new random schedules, making copies of high quality schedules 
+      in the present generation, applies the mutation operator to schedules, or uses the
+      crossover operator on 2 schedules. When performing mutation or crossover, the 
+      schedules are selected with a probability proportional to their fitness, allowing 
+      the schedules with high quality tobe selected more frequently and make more 
+      "offspring".
+
+      The free parameters in our GA are the size of the population, number of
+      generations to simulate, and proportions of new generations toi be made by the copy, 
+      random, mutation, and crossover methods.
+
+    \subsubsection{Representation}
+      We use the naive representation of schedules. A schedule contains lists of size $|C|$
+      containing the room in which each course's exam is to occur and the timing
+      (corresponding to a pair or integers representing the day and time) of each course's
+      exam.
+
     \subsubsection{Mutation Operator}
-      [\ldots]
+      The mutation operator takes a probability and a schedule. We loop over the courses,
+      and, with the given probability, each is or is not set to a random room at a random time.
+
     \subsubsection{Crossover Operator}
-      [\ldots]
+      Given two schedules, the crossover operator loops over the courses, and chooses at
+      random which schedule to take the room and timing from for that course. This results 
+      in an interleaving of the original courses' data.
+
+      There is, unfortunately, no good reason to expect that the result of applying 
+      crossover to two relatively high quality schedules will produce a high quality 
+      schedule, since the quality of schedules can be very subtle, as they are a more 
+      global property of the schedule than this process captures. The crossover of two 
+      schedules where no student or room has a timing conflict is unlikely to result in 
+      a schedule with much more than average quality. There is no obvious reason to think 
+      that this sort of procedure will produce schedules with better than random fitness. 
+      Luckily, the schedules produced in practice are better than random.
+
+      If a different representation were chosen, such that the mutated schedules were
+      different from, but likely to have quality similar to their parents, the GA would 
+      likely be able to perform much better than it does, but we could not discover such a
+      representation and are satisfied with this method for present purposes.
+
   \subsection{Other Search Algorithms}
@@
     \subsubsection{Mutate Search}
-      Mutate Search starts by generating a random schedule. It modifies this schedule (by
-      using the same mutate operator made for the genetic algorithm) by 1\% and rolls back if
-      the mutation didn't increase the schedule's fitness. This operation is repeated until it has
-      performed as many evaluations as it is allowed to.
-      This algorithm is a variation of simulated annealing, where instead of accepting a change
-      through a probabilistic function, it always accepts a change that leads to a higher fitness.
+      Mutate Search starts by generating a random schedule. It makes an alternative schedule
+      (by using the same mutate operator made for the genetic algorithm) with 1\% of the
+      data randomized and chooses which of the schedules is better. This operation is
+      repeated until it has performed as many evaluations as it is allowed to.
+
+      This method can be changed to true simulated annealing by varying the degree of
+      modification made to the current best schedule with the number of evaluations
+      remaining. Not being the focus of our investigation, little effort was made to 
+      optimize the parameters. 1\% does seem to give better performance than other nearby 
+      values, though we did not collect data to substantiate this. Finding ideal parameters 
+      for simulated annealing on this problem would make for a project-sized inquiry in itself.
+
+  \subsection{Options for the Genetic Algorithm}
+    \subsubsection{Scheduling Problem Instances}
+      We generated 2 sets of 3 scheduling problem instances to test the algorithms above with. 
+      The package ScheduleInstanceGenerator in our code repository manages the generation of 
+      random problem instances with many free parameters to specify characteristics of the 
+      output instance.
+
+      The following table characterized the problem instances that we used:
+
+      \begin{tabular}{|c|c|c|c|c|c|c|}
+      \hline  & Tiny & Medium & Large & Tiny /w Cor. & Med /w Cor. & Large /w Cor. \\ 
+      \hline d & 7 & 10 & 7 & 7 & 10 & 7 \\ 
+      \hline t & 5 & 8 & 8 & 5 & 8 & 8 \\ 
+      \hline $|C|$ & 20 & 200 & 603 & 20 & 200 & 600 \\ 
+      \hline $|R|$ & 4 & 10 & 43 & 4 & 10 & 43 \\ 
+      \hline $|S|$ & 50 & 300 & 21945 & 50 & 300 & 21945 \\ 
+      \hline $|L_s|$ & 1-5 & 1-6 & 1-5 & 2-6 & 2-6 & 2-6 \\ 
+      \hline \# Course Types & N/A & N/A & N/A & 4 & 10 & 20 \\ 
+      \hline 
+	\end{tabular}
+
+	  The full input files for these instances are available in our repository in the 
+	  directory "Instance Input Files".
+
+    \subsubsection{Selecting Parameters for the GA}
+      [\ldots]
+  \subsection{Comparison Between the Different Searches}
+    [\ldots]
 
@@
     algorithms will very quickly tend to reject all such schedules.
-  \subsection{The Genetic Algorithm's Parameters}
-    [\ldots]
-  \subsection{Comparison Between the Different Searches}
-    [\ldots]
 
@@
   is a large data set. But it may not be the best solution for this problem, since
-  the MutateSearch performs way better.
+  the Mutate Search performs way better.
 
-  %\cite{website:jgap} 
-  %\cite{website:jenes} 
-  %\cite{website:coderepos} 
 \bibliographystyle{aaai}

### 2009-11-30 Paper/paper.tex
Paper updated, finally happy with the table

--- 2009-11-24 trunk/Paper/paper.tex
@@
 \usepackage{courier}
+\usepackage{url}
+\usepackage[table]{xcolor}
 % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\begin{abstract}
+  We examine the exam scheduling problem with soft constraints for problems of varying
+  sizes using a genetic algorithm, random search, hillclimbing, and a variant of simulated
+  annealing. This is done in the context of maximizing the fitness of a schedule, where the
+  fitness is relatively computationally expensive to compute for any schedule. The genetic
+  algorithm is discussed in detail. We consider selecting parameters for the GA that
+  maximize the quality of the schedule found after a constant number of evaluations. The
+  GA, with a naive crossover operator, performs much better than the random search, but
+  simulated annealing is far superior in this search space.
+\end{abstract}
+
 \section{The Problem}
-Description of our problem
+  The problem we investigate is a sort of scheduling problem. 
+  We are attempting to find the most satisfactory choice of when
+  and where to hold exams, given a background of student course loads.
+
+  An instance of this scheduling problem consists of a number of days
+  on which exams can be scheduled ($d$), the number of time slots in which
+  an exam can be scheduled on any day ($t$), a set of rooms ($R$), a set of
+  courses ($C$), and a set of students ($S$), each of whom ($s$) has a
+  particular course load, some subset of the courses ($L_s$).
+
+  A schedule, then, is a mapping from courses to rooms at times, 
+  which we can express like this:
+
+  \[ C \rightarrow \left\{(r,\ a,\ b)\ |\ r\in R,\ a\in \{1..d\},\ b\in \{1..t\}\right\} \]
+
+  The total number of different possible schedules in any problem instance is
+  $|C|^{|R|dt}$ which is typically far too large for brute force search.
+  For example, one instance that we examine (scheduling the fall semester
+  exams at the University of Toronto in 2009) involves 603 courses, 7 days,
+  8 times, 43 rooms, and 21945 students. This instance admits approximately 
+  $10^{6695}$ different possible schedules.
+
+  Typically, with relatively loose constraints on the number of rooms and 
+  particular schedules of students, it is not difficult to find some consistent
+  schedule; that is, one where no student is asked to write 2 exams
+  simultaneously and no room has 2 exams occur in it simultaneously.
+
+  Rather than worry about hard constraints, we prefer a framework of soft 
+  constraints, where we try to find a schedule that makes the students and 
+  invigilators happiest overall, allowing the possibility that some student 
+  or room is left with an impossible exam schedule, which can, in the real 
+  world, be dealt with on an individual basis.
+
+  The timetable that a student $s \in S$ has under some particular schedule 
+  $K$ may be represented in this way:
+
+  \[ TT_s(K) = \left\{(a, b)\ |\ \exists r \in R, \exists c \in L_s, K(c)=(r, a, b) \right\} \]
+
+  which may, in general, be a multiset.
+
+  Similarly, the timetable for a room $r \in R$ is
+
+  \[ TT_r(K)=\left\{(a, b)\ |\ \exists c\in C,\ K(c)=(r, a, b)\right\} \]
+
+  We define two quality functions, mapping schedules to real numbers in [0,1],
+  one for students and one for rooms. These are meant to capture how much they
+  "like" the schedule under consideration. (For instance, a student will rate
+  poorly any schedule where she has 2 consecutive exams, and very poorly any 
+  schedule which expects her to take two exams at the same time.)
+
+  \[ Q_s(K)=q_{student}\left(TT_s(K)\right):\textit{ Schedules}\rightarrow [0,\ 1] \]
+  \[ Q_r(K)=q_{room}\left(TT_r(K)\right):\textit{ Schedules}\rightarrow [0,\ 1] \]
+
+  We will consider later how to define these functions in a reasonable way.
+
+  The quality of a schedule, then, is given by
+
+  \[ Q(K)= \frac{\left(w_s\left(\sum_{s \in S}{\frac{Q_s(K)}{|S|}}\right) +
+     w_r\left(\sum_{r\in R}{\frac{Q_r(K)}{|R|}} \right)\right)}{(w_s+w_r)} \]
+
+  where $w_s$ and $w_r$ are the weightings we can use to indicate that we 
+  care somewhat more about students preferences than rooms, giving a quality
+  (or fitness, in the terminology of genetic algorithms) in the range [0,1] for
+  any schedule.
+
+  Evaluating this function Q can be computationally expensive when $|S|$ and $|R|$
+  are large. Our work examines how to find a schedule with relatively high fitness
+  considering that we will want to do this with as few evaluations of Q as possible.
+
+  We examine, first and foremost, a genetic algorithm approach to solving this problem,
+  but also touch on a few other methods.
 
 \section{Approach and Implementation}
-Tools, algorithms, APIs, etc. that we used.
+  \subsection{Differences From Natural Exam Scheduling Problems}
+    The problem we have constructed here is a compromise between the most
+    mathematically pure problem with the same sort of properties (which might ignore the
+    rooms and related constraints completely) and the problem that must be solved in the
+    real world when scheduling thousands of students, which has some important differences
+    from this problem.
+
+  \subsection{Tools used}
+    The program is implemented in Java, to generate possible schedules and to calculate
+    their fitness. A Mathematica program is also written to plot graphs, given the data
+    generated by the Java program. Additionally, we use Google code 
+    subversion\footnote{\url{http://code.google.com/p/csc384-genetic-algorithms-project/}}
+    and Google doc to share the work.
+
+  \subsection{Fitness Function}
+    In order to assign each schedule a fitness, we need to define the functions 
+    $q_{student}$ and $q_{room}$. They need to capture the idea of determining how
+    much each student or room likes their given timetable. We do this by calculating 
+    the penalty incurred by a room, which is a measure of it having properties that 
+    are unpleasant to the student or room, and then using the formula
+
+    \[ q(TT)=\frac{1}{1+penalty_{TT}} \]
+
+    This function will be in the desired range or [0,1].
+
+    For students, we accrue a penalty of 50 for every time that is repeated in a 
+    timetable, 5 when there are exams in 2 consecutive times, 3 when 2 exams happen 
+    on the same day, and 0.5 when the student has exams on consecutive days. This 
+    accords reasonably well with real students' hopes for their exam timetables.
+
+    Rooms have the same penalty of 50 when it is used by two exams at the same time, 
+    5 if it hosts exams in consecutive time slots (reasoning that the invigilators 
+    need time to let the students out and in and prepare for the next exam) and a 
+    penalty of 0.5 when a room is empty for an entire day between uses (reasoning 
+    that the room requires some amount of modifications to be used for exams, and 
+    it is difficult to use it for other purposes if it will still be used for exams 
+    soon.)
+
+    The details of the function we used are essentially arbitrary.
+
+  \subsection{Genetic Algorithm Search}
+    Although there are many readily available open source genetic algorithm software
+    packages available (see JGAP \cite{website:JGAP}, Jenes\cite{website:jenes}) we 
+    decided to write a simple, yet general, genetic algorithm package, in Java, that
+    can interface with data types designed by others, so long as they implement a few
+    necessary methods. The package we wrote is more than adequate for investigating
+    the scheduling problem under consideration here.
+
+    \subsubsection{General Description}
+      The genetic algorithm maintains a population of schedules and constructs subsequent
+      generations by adding new random schedules, making copies of high quality schedules 
+      in the present generation, applies the mutation operator to schedules, or uses the
+      crossover operator on 2 schedules. When performing mutation or crossover, the 
+      schedules are selected with a probability proportional to their fitness, allowing 
+      the schedules with high quality to be selected more frequently and make more 
+      "offspring".
+
+      The free parameters in our GA are the size of the population, number of
+      generations to simulate, and proportions of new generations to be made by the copy, 
+      random, mutation, and crossover methods.
+
+    \subsubsection{Representation}
+      We use the naive representation of schedules. A schedule contains lists of size $|C|$
+      containing the room in which each course's exam is to occur and the timing
+      (corresponding to a pair or integers representing the day and time) of each course's
+      exam.
+
+    \subsubsection{Mutation Operator}
+      The mutation operator takes a probability and a schedule. We loop over the courses,
+      and, with the given probability, each is or is not set to a random room at a random time.
+
+    \subsubsection{Crossover Operator}
+      Given two schedules, the crossover operator loops over the courses, and chooses at
+      random which schedule to take the room and timing from for that course. This results 
+      in an interleaving of the original courses' data.
+
+      There is, unfortunately, no good reason to expect that the result of applying 
+      crossover to two relatively high quality schedules will produce a high quality 
+      schedule, since the quality of schedules can be very subtle, as they are a more 
+      global property of the schedule than this process captures. The crossover of two 
+      schedules where no student or room has a timing conflict is unlikely to result in 
+      a schedule with much more than average quality. There is no obvious reason to think 
+      that this sort of procedure will produce schedules with better than random fitness. 
+      Luckily, the schedules produced in practice are better than random.
+
+      If a different representation were chosen, such that the mutated schedules were
+      different from, but likely to have quality similar to their parents, the GA would 
+      likely be able to perform much better than it does, but we could not discover such a
+      representation and are satisfied with this method for present purposes.
+
+  \subsection{Other Search Algorithms}
+    \subsubsection{Random Search}
+      The random search method simply generates random schedules, evaluates their fitness,
+      and remembers the one that it has seen that has highest fitness. This method is meant
+      as a baseline against which to compare the other methods. If a search method doesn't
+      perform significantly better than random search with the same number of fitness
+      evaluations, it is useless.
+
+    \subsubsection{Hillclimbing Search}
+      The hillclimbing search that we have implemented examines every schedule in the
+      neighbourhood of a given starting schedule, and then continues from the one with
+      highest fitness, terminating when it has performed as many evaluations as it is allowed
+      to, or it has found a schedule whose neighbour are all equal or inferior to it.
+
+      The neighbouring schedules are those where only one course meets in either a different
+      room, on a different day, or at a different time. This leads to a total of
+      $ |C|(d+t+|R|-3) $ schedules in the neighbourhood of any given schedule.
+
+      It can be seen that this method is completely deterministic given a starting schedule,
+      and will always return a local maximum
+
+    \subsubsection{Mutate Search}
+      Mutate Search starts by generating a random schedule. It makes an alternative schedule
+      (by using the same mutate operator made for the genetic algorithm) with 1\% of the
+      data randomized and chooses which of the schedules is better. This operation is
+      repeated until it has performed as many evaluations as it is allowed to.
+
+      This method can be changed to true simulated annealing by varying the degree of
+      modification made to the current best schedule with the number of evaluations
+      remaining. Not being the focus of our investigation, little effort was made to 
+      optimize the parameters. 1\% does seem to give better performance than other nearby 
+      values, though we did not collect data to substantiate this. Finding ideal parameters 
+      for simulated annealing on this problem would make for a project-sized inquiry in itself.
+
+  \subsection{Options for the Genetic Algorithm}
+    \subsubsection{Scheduling Problem Instances}
+      We generated 2 sets of 3 scheduling problem instances to test the algorithms above with. 
+      The package \emph{ScheduleInstanceGenerator} in our code repository manages the generation of 
+      random problem instances with many free parameters to specify characteristics of the 
+      output instance.
+
+      The following table characterized the problem instances that we used:
+
+      \definecolor{light-gray}{gray}{0.87}
+      \rowcolors{1}{light-gray}{white}
+      \begin{center}
+      \begin{small}
+      \begin{tabular}{*{7}{c}}
+      \hline 
+         		& T 	& M 	& L 	& $T_{Cor}$ & $M_{Cor}$ & $L_{Cor}$ \\ 
+      \hline \hline
+       d 		& 7 	& 10	& 7 	& 7 	& 10 	& 7 	\\ 
+       t 		& 5		& 8 	& 8 	& 5 	& 8 	& 8 	\\ 
+       $|C|$	& 20 	& 200 	& 603 	& 20 	& 200 	& 600 	\\ 
+       $|R|$	& 4 	& 10 	& 43 	& 4 	& 10 	& 43	\\ 
+       $|S|$	& 50 	& 300 	& 21945 & 50 	& 300 	& 21945	\\ 
+       $|L_s|$	& 1-5 	& 1-6 	& 1-5 	& 2-6 	& 2-6 	& 2-6 	\\ 
+       $|T|$	& N/A 	& N/A 	& N/A 	& 4 	& 10 	& 20 	\\ 
+       s		& $10^{182}$ & $10^{1840}$ & $10^{6695}$ & $10^{182}$ & $10^{1840}$ & $10^{6689}$ \\
+      \hline 
+	  \end{tabular} 
+	  \end{small}
+	  \end{center}
+
+	  The large data set was created with parameters inspired from the University of 
+	  Toronto's Faculty of Arts and Science \cite{website:uoftstats,website:artscistats}.
+	  The full input files for these instances are available in our repository in the 
+	  directory \emph{Instance Input Files}.
+
+    \subsubsection{Selecting Parameters for the GA}
+      [\ldots]
+
+  \subsection{Comparison Between the Different Searches}
+    [\ldots]
 
 \section{Evaluation}
-How we determined the success of our project. Did we solve it? Effectively? Labelled images, graphs and/or tables can be useful.
+  \subsection{Success}
+    [\ldots]
+    Seeing as our fitness function gives a very high penalty for schedules that have direct
+    conflicts (either a student with two exams at the same moment, or two different exams
+    in a same room at the same moment), if such schedules exists given the input data, 
+    our algorithms will tend to very quickly reject them.
 
-% \bibliography{Bibliography-File}
-% \bibliographystyle{aaai}
+\section{Conclusion}
+  [\ldots]
+  We come up with a conclusion that in this particular problem, genetic algorithm
+  is better than random search, and it is better than hill-climbing if the input 
+  is a large data set. But it may not be the best solution for this problem, since
+  the Mutate Search performs way better.
+
+\bibliographystyle{aaai}
+\bibliography{paperBib}
 \end{document}

### 2009-12-01 Paper/paper.tex
Paper update; first graph inserted

@@
 \usepackage[table]{xcolor}
+\usepackage{graphicx}
 % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     Although there are many readily available open source genetic algorithm software
-    packages available (see JGAP \cite{website:JGAP}, Jenes\cite{website:jenes}) we 
+    packages available (see JGAP, Jenes) we 
     decided to write a simple, yet general, genetic algorithm package, in Java, that
@@
       We generated 2 sets of 3 scheduling problem instances to test the algorithms above with. 
-      The package \emph{ScheduleInstanceGenerator} in our code repository manages the generation of 
+      The package \texttt{ScheduleInstanceGenerator} in our code repository manages the generation of 
       random problem instances with many free parameters to specify characteristics of the 
       output instance.
+
+      The first set of instances were generated with the parameters in the table below. 
+      The rooms and courses and students were given arbitrary names and the courses that each 
+      student attended were selected with uniform probability from the complete set of courses.
+
+      The second set was generated with a sort of correlation between the courses that each student 
+      took, rendering the dataset more true to the sort of data that would be found in real-world 
+      instances of the exam scheduling problem. The set of courses that each student takes was 
+      selected in the following way. Each student was assigned, with uniform probability, a \emph{major} 
+      and a random number of courses between 2 and 6. Each course was then selected, with probability 
+      0.7, from among the $|S|/\#Majors$ courses in their major, or, with probability 0.3, uniformly from 
+      the set of all courses. We can expect that each student takes 70\% of their courses from a small 
+      subset of the courses, representing the field that they specialize in, and 30\% from other fields. 
+      This simulates the way in which we expect student bodies to choose their courses, though the true 
+      data is likely to exhibit far more structure, due to program requirements, notorious professors, etc.
+
+      The 3 sizes vary from manageable to large. The smallest could potentially be solved exactly, 
+      with methods not under consideration here, while the others have search spaces much too large 
+      for any hope of an exact solution. For the largest, which simulates a UofT exam schedule, a 
+      single evaluation of the fitness function took 0.2 seconds on the fastest machine owned by any 
+      of the authors.
 
@@
       \hline 
-         		& T 	& M 	& L 	& $T_{Cor}$ & $M_{Cor}$ & $L_{Cor}$ \\ 
+      	& S 	& M 	& L 	& $S_{Cor}$ & $M_{Cor}$ & $L_{Cor}$ \\ 
       \hline \hline
@@
        $|L_s|$	& 1-5 	& 1-6 	& 1-5 	& 2-6 	& 2-6 	& 2-6 	\\ 
-       $|T|$	& N/A 	& N/A 	& N/A 	& 4 	& 10 	& 20 	\\ 
-       s		& $10^{182}$ & $10^{1840}$ & $10^{6695}$ & $10^{182}$ & $10^{1840}$ & $10^{6689}$ \\
+       \begin{tiny}$\#$Maj\end{tiny}	& N/A 	& N/A 	& N/A 	& 4 	& 10 	& 20 	\\ 
+       \begin{tiny}\textsc{sss}\end{tiny}	& $10^{182}$ & $10^{1840}$ & $10^{6695}$ & $10^{182}$ & $10^{1840}$ & $10^{6689}$ \\
       \hline 
@@
 	  The large data set was created with parameters inspired from the University of 
-	  Toronto's Faculty of Arts and Science \cite{website:uoftstats,website:artscistats}.
+	  Toronto's Faculty of Arts and Science.
 	  The full input files for these instances are available in our repository in the 
@@
     [\ldots]
+    \begin{figure}[h!]
+  	  \centering
+  	  \setlength\fboxsep{0pt}
+	  \setlength\fboxrule{0.5pt}
+	  \fbox{\includegraphics[width=8.35cm]{SMALL_FullAverageStats}}
+      \caption{Search statistics on Small data set}
+	\end{figure}
 
@@
 
+  \nocite{*}
 \bibliographystyle{aaai}

### 2009-12-01 Paper/paper.tex
Updated graph on paper

@@
 \usepackage{graphicx}
+\usepackage{epstopdf}
 % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
        \begin{tiny}$\#$Maj\end{tiny}	& N/A 	& N/A 	& N/A 	& 4 	& 10 	& 20 	\\ 
-       \begin{tiny}\textsc{sss}\end{tiny}	& $10^{182}$ & $10^{1840}$ & $10^{6695}$ & $10^{182}$ & $10^{1840}$ & $10^{6689}$ \\
+       \textsc{sss}	& $10^{182}$ & $10^{1840}$ & $10^{6695}$ & $10^{182}$ & $10^{1840}$ & $10^{6689}$ \\
       \hline 
@@
     \subsubsection{Selecting Parameters for the GA}
-      [\ldots]
+      In order to select parameters for the genetic algorithm (that is, the population, 
+      number of generations to simulate, proportions of copies, randoms, mutations, and 
+      crossovers for the next generation) we performed tests using the small uncorrelated 
+      problem instance.
+
+      \paragraph{Generation Parameters}
+        We undertook a search of the space of parameters for forming new generations.
+
+        In the graph below, each data point represents the average of 5 trials. To make 
+        the scale clearer, the colour on the surface corresponds to the fitness, where 
+        the highest fitnesses are red. The number of copies is fixed at 1, and the number 
+        of mutations (m) and crossovers (c) are shown on the horizontal axes. The total 
+        size of the population is fixed, so the number of random individuals used at each 
+        point is 60?-(m+c). This corresponds to the distance from the diagonal.
+
+        [GRAPH]
+
+        The graph above indicates that for a fixed number of evaluations, best results are 
+        achieved by selecting parameters that have the majority of individuals in the next 
+        generation be generated by crossovers of the most fit schedules in the previous 
+        generation. In order to achieve the very highest fitness, a few randoms or mutations 
+        should also be incorporated.
+
+        This test was done for only the smallest problem instance, and is infeasible for the 
+        larger instances. We chose generation proportion parameters to approximate the optimal 
+        values seen here, under the assumption that the larger problems exhibited this sort of 
+        preference.
+
+      \paragraph{Size Parameters}
+        Blah blah blaj. We investigated the population size and
+
+        [Insert fancy 3D graph here = WIN]
 
@@
 	  \setlength\fboxrule{0.5pt}
-	  \fbox{\includegraphics[width=8.35cm]{SMALL_FullAverageStats}}
-      \caption{Search statistics on Small data set}
+	  \fbox{\includegraphics[width=8.35cm]{HUGE-TYPE2_FullAverageStats}}
+      \caption{Search comparison on a large data set}
 	\end{figure}

### 2009-12-01 trunk/Paper/paper.tex
Updated table in paper, not perfect, but good enough for the moment.
Paper update, 3D graphs in it

@@
 \usepackage{url}
+\usepackage[table]{xcolor}
+\usepackage{graphicx}
+\usepackage{epstopdf}
+\usepackage{float}
 % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
       schedules are selected with a probability proportional to their fitness, allowing 
-      the schedules with high quality tobe selected more frequently and make more 
+      the schedules with high quality to be selected more frequently and make more 
       "offspring".
@@
       The free parameters in our GA are the size of the population, number of
-      generations to simulate, and proportions of new generations toi be made by the copy, 
+      generations to simulate, and proportions of new generations to be made by the copy, 
       random, mutation, and crossover methods.
@@
       We generated 2 sets of 3 scheduling problem instances to test the algorithms above with. 
-      The package ScheduleInstanceGenerator in our code repository manages the generation of 
+      The package \texttt{ScheduleInstanceGenerator} in our code repository manages the generation of 
       random problem instances with many free parameters to specify characteristics of the 
@@
 
+      The first set of instances were generated with the parameters in the table below. 
+      The rooms and courses and students were given arbitrary names and the courses that each 
+      student attended were selected with uniform probability from the complete set of courses.
+
+      The second set was generated with a sort of correlation between the courses that each student 
+      took, rendering the dataset more true to the sort of data that would be found in real-world 
+      instances of the exam scheduling problem. The set of courses that each student takes was 
+      selected in the following way. Each student was assigned, with uniform probability, a \emph{major} 
+      and a random number of courses between 2 and 6. Each course was then selected, with probability 
+      0.7, from among the $|S|/\#Majors$ courses in their major, or, with probability 0.3, uniformly from 
+      the set of all courses. We can expect that each student takes 70\% of their courses from a small 
+      subset of the courses, representing the field that they specialize in, and 30\% from other fields. 
+      This simulates the way in which we expect student bodies to choose their courses, though the true 
+      data is likely to exhibit far more structure, due to program requirements, notorious professors, etc.
+
+      The 3 sizes vary from manageable to large. The smallest could potentially be solved exactly, 
+      with methods not under consideration here, while the others have search spaces much too large 
+      for any hope of an exact solution. For the largest, which simulates a UofT exam schedule, a 
+      single evaluation of the fitness function took 0.2 seconds on the fastest machine owned by any 
+      of the authors.
+
       The following table characterized the problem instances that we used:
 
-      \begin{tabular}{|c|c|c|c|c|c|c|}
-      \hline  & Tiny & Medium & Large & Tiny /w Cor. & Med /w Cor. & Large /w Cor. \\ 
-      \hline d & 7 & 10 & 7 & 7 & 10 & 7 \\ 
-      \hline t & 5 & 8 & 8 & 5 & 8 & 8 \\ 
-      \hline $|C|$ & 20 & 200 & 603 & 20 & 200 & 600 \\ 
-      \hline $|R|$ & 4 & 10 & 43 & 4 & 10 & 43 \\ 
-      \hline $|S|$ & 50 & 300 & 21945 & 50 & 300 & 21945 \\ 
-      \hline $|L_s|$ & 1-5 & 1-6 & 1-5 & 2-6 & 2-6 & 2-6 \\ 
-      \hline \# Course Types & N/A & N/A & N/A & 4 & 10 & 20 \\ 
+      \definecolor{light-gray}{gray}{0.87}
+      \rowcolors{1}{light-gray}{white}
+      \begin{center}
+      \begin{small}
+      \begin{tabular}{*{7}{c}}
+      \hline 
+      	& S 	& M 	& L 	& $S_{Cor}$ & $M_{Cor}$ & $L_{Cor}$ \\ 
+      \hline \hline
+       d 		& 7 	& 10	& 7 	& 7 	& 10 	& 7 	\\ 
+       t 		& 5		& 8 	& 8 	& 5 	& 8 	& 8 	\\ 
+       $|C|$	& 20 	& 200 	& 603 	& 20 	& 200 	& 600 	\\ 
+       $|R|$	& 4 	& 10 	& 43 	& 4 	& 10 	& 43	\\ 
+       $|S|$	& 50 	& 300 	& 21945 & 50 	& 300 	& 21945	\\ 
+       $|L_s|$	& 1-5 	& 1-6 	& 1-5 	& 2-6 	& 2-6 	& 2-6 	\\ 
+       \begin{tiny}$\#$Maj\end{tiny}	& N/A 	& N/A 	& N/A 	& 4 	& 10 	& 20 	\\ 
+       \textsc{sss}	& $10^{182}$ & $10^{1840}$ & $10^{6695}$ & $10^{182}$ & $10^{1840}$ & $10^{6689}$ \\
       \hline 
 	  \end{tabular} 
+	  \end{small}
+	  \end{center}
 
+	  The large data set was created with parameters inspired from the University of 
+	  Toronto's Faculty of Arts and Science.
 	  The full input files for these instances are available in our repository in the 
-	  directory "Instance Input Files".
+	  directory \emph{Instance Input Files}.
 
     \subsubsection{Selecting Parameters for the GA}
-      [\ldots]
+      In order to select parameters for the genetic algorithm (that is, the population, 
+      number of generations to simulate, proportions of copies, randoms, mutations, and 
+      crossovers for the next generation) we performed tests using the small uncorrelated 
+      problem instance.
+
+      \paragraph{Generation Parameters}
+        We undertook a search of the space of parameters for forming new generations.
+
+        In the graph below, each data point represents the average of 5 trials. To make 
+        the scale clearer, the colour on the surface corresponds to the fitness, where 
+        the highest fitnesses are red. The number of copies is fixed at 1, and the number 
+        of mutations (m) and crossovers (c) are shown on the horizontal axes. The total 
+        size of the population is fixed, so the number of random individuals used at each 
+        point is 60?-(m+c). This corresponds to the distance from the diagonal.
+
+        \begin{figure}[H]
+  		  \centering
+  		  \includegraphics[width=8cm]{GAParametersData}
+  	      \caption{Effectiveness of the GA's proportion parameters}
+  	      \label{graph:GAParam1}
+		\end{figure}
+
+        The graph above indicates that for a fixed number of evaluations, best results are 
+        achieved by selecting parameters that have the majority of individuals in the next 
+        generation be generated by crossovers of the most fit schedules in the previous 
+        generation. In order to achieve the very highest fitness, a few randoms or mutations 
+        should also be incorporated.
+
+        This test was done for only the smallest problem instance, and is infeasible for the 
+        larger instances. We chose generation proportion parameters to approximate the optimal 
+        values seen here, under the assumption that the larger problems exhibited this sort of 
+        preference.
+
+      \paragraph{Size Parameters}
+        Blah blah blaj. We investigated the population size and
+
+        \begin{figure}[H]
+  		  \centering
+  		  \includegraphics[width=8cm]{MEDIUM-GA_Size_Data}
+  	      \caption{Effectiveness of the GA's population parameters}
+  	      \label{graph:GAParam2}
+		\end{figure}
+
   \subsection{Comparison Between the Different Searches}
     [\ldots]
+    \begin{figure}[h!]
+  	  \centering
+  	  \setlength\fboxsep{0pt}
+	  \setlength\fboxrule{0.5pt}
+	  \fbox{\includegraphics[width=8.35cm]{HUGE-TYPE2_FullAverageStats}}
+      \caption{Search comparison on a large data set}
+	\end{figure}
 
@@
     conflicts (either a student with two exams at the same moment, or two different exams
-    in a same room at the same moment), if such a schedule exists given the input data, our
-    algorithms will very quickly tend to reject all such schedules.
+    in a same room at the same moment), if such schedules exists given the input data, 
+    our algorithms will tend to very quickly reject them.
 
 \section{Conclusion}
+  [\ldots]
   We come up with a conclusion that in this particular problem, genetic algorithm

### 2009-12-01 trunk/Paper/paper.tex
Paper update, close to finish, few changes from place to place, temporary Lorem ipsum to fill the ending

@@
     real world when scheduling thousands of students, which has some important differences
-    from this problem.
+    from this problem. For example, we don't deal with constraints such as room capacity
+    or exams with varying length.
 
   \subsection{Tools used}
-    The program is implemented in Java, to generate possible schedules and to calculate
-    their fitness. A Mathematica program is also written to plot graphs, given the data
-    generated by the Java program. Additionally, we use Google code 
-    subversion\footnote{\url{http://code.google.com/p/csc384-genetic-algorithms-project/}}
-    and Google doc to share the work.
+    Our application was entirely written in Java. We also wrote another Java application 
+    to generate problem instance files for testing our algorithms. We used Mathematica 
+    to plot all of our graphs.  Additionally, we use Google code's source code
+    repository\footnote{\url{http://code.google.com/p/csc384-genetic-algorithms-project/}} 
+    and Google docs to collaborate on this report.
 
@@
 
-    The details of the function we used are essentially arbitrary.
+    The details of the function we used are essentially arbitrary, and we expect 
+    similar results would be attained for any other penalty-based fitness function 
+    modelling the preferences of students and requirements of rooms.
 
@@
     Although there are many readily available open source genetic algorithm software
-    packages available (see JGAP, Jenes) we 
-    decided to write a simple, yet general, genetic algorithm package, in Java, that
-    can interface with data types designed by others, so long as they implement a few
-    necessary methods. The package we wrote is more than adequate for investigating
-    the scheduling problem under consideration here.
+    packages available (see JGAP, Jenes) we decided to write a simple, yet general, 
+    genetic algorithm package, in Java, that can interface with data types designed 
+    by others, so long as they implement a few necessary methods. The package we wrote 
+    is more than adequate for investigating the scheduling problem under consideration here.
 
@@
       the schedules with high quality to be selected more frequently and make more 
-      "offspring".
+      ``offspring''.
 
@@
       exam.
+
+      We will see that this representation is far from ideal, but there is no clear alternative.
 
@@
 
-      The following table characterized the problem instances that we used:
-
+      The following table characterized the problem instances that we used 
+      (with \textsc{SSS} being an approximation of the search space size):      
       \definecolor{light-gray}{gray}{0.87}
@@
        \begin{tiny}$\#$Maj\end{tiny}	& N/A 	& N/A 	& N/A 	& 4 	& 10 	& 20 	\\ 
-       \textsc{sss}	& $10^{182}$ & $10^{1840}$ & $10^{6695}$ & $10^{182}$ & $10^{1840}$ & $10^{6689}$ \\
+       \textsc{SSS}	& $10^{182}$ & $10^{1840}$ & $10^{6695}$ & $10^{182}$ & $10^{1840}$ & $10^{6689}$ \\
       \hline 
@@
 	  Toronto's Faculty of Arts and Science.
-	  The full input files for these instances are available in our repository in the 
-	  directory \emph{Instance Input Files}.
 
@@
         the scale clearer, the colour on the surface corresponds to the fitness, where 
-        the highest fitnesses are red. The number of copies is fixed at 1, and the number 
-        of mutations (m) and crossovers (c) are shown on the horizontal axes. The total 
-        size of the population is fixed, so the number of random individuals used at each 
-        point is 60?-(m+c). This corresponds to the distance from the diagonal.
+        the highest fitnesses are red. The number of copies is fixed at 1, and the proportions 
+        of mutations and crossovers are shown on the horizontal axes.
 
@@
   		  \centering
-  		  \includegraphics[width=8cm]{GAParametersData}
+  		  \includegraphics[width=6cm]{GAParametersData}
   	      \caption{Effectiveness of the GA's proportion parameters}
@@
       \paragraph{Size Parameters}
-        Blah blah blaj. We investigated the population size and
+       [\ldots]  Duis accumsan sagittis velit vel consectetur.
+Nullam sed ligula nisl.
+Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.
+Morbi quis neque eu justo accumsan auctor.
+Cras a risus ac est aliquam malesuada quis ac tellus.
+Aliquam nunc diam, ultrices et cursus ac, ornare mollis tellus.
+Ut venenatis lacus nec risus eleifend placerat.
 
@@
 
+		[\ldots]  Praesent eros nulla, ullamcorper non euismod in, pretium et nisi.
+Nam eu augue ut lacus dignissim gravida.
+Vivamus euismod convallis lorem, bibendum pellentesque arcu pulvinar a.
+Phasellus suscipit vehicula nunc non laoreet.
+
   \subsection{Comparison Between the Different Searches}
-    [\ldots]
-    \begin{figure}[h!]
+    [\ldots] Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum fermentum vehicula tellus, id varius enim cursus nec.
+In auctor lacus ac diam iaculis id fringilla est tincidunt.
+Sed tincidunt dictum dolor non dapibus.
+Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Phasellus bibendum nisl quis felis accumsan vel hendrerit nulla gravida.
+Fusce convallis velit eu nisi pellentesque quis sagittis dolor imperdiet.
+Sed magna velit, convallis in mattis vel, varius vel ante.
+Nunc euismod aliquam lacus vitae consectetur.
+
+    \begin{figure}[H]
   	  \centering
@@
 	  \setlength\fboxrule{0.5pt}
-	  \fbox{\includegraphics[width=8.35cm]{HUGE-TYPE2_FullAverageStats}}
-      \caption{Search comparison on a large data set}
+	  \fbox{\includegraphics[width=8cm]{HUGE-TYPE2_FullAverageStats}}
+      \caption{Comparison of the different searches on a large data set}
 	\end{figure}
 
+	[\ldots]Vivamus vehicula metus nibh.
+Morbi id ante dolor, eget pretium justo.
+Phasellus sit amet arcu vitae eros malesuada ornare.
+Proin vulputate volutpat eros, a laoreet ipsum lacinia vel.
+Vivamus commodo odio non tellus convallis pharetra.
+Cras varius fringilla tempor.
+Sed mollis, augue quis euismod pellentesque, turpis mi convallis purus, sed bibendum eros lacus volutpat felis.
+Quisque velit lectus, dapibus eget tincidunt sed, tincidunt a lectus.
+Phasellus vel luctus eros.
+Vestibulum vitae quam mi.
+Nullam posuere aliquam magna vitae bibendum.
+
 \section{Evaluation}
-  \subsection{Success}
-    [\ldots]
-    Seeing as our fitness function gives a very high penalty for schedules that have direct
-    conflicts (either a student with two exams at the same moment, or two different exams
-    in a same room at the same moment), if such schedules exists given the input data, 
-    our algorithms will tend to very quickly reject them.
 
 \section{Conclusion}
-  [\ldots]
-  We come up with a conclusion that in this particular problem, genetic algorithm
-  is better than random search, and it is better than hill-climbing if the input 
-  is a large data set. But it may not be the best solution for this problem, since
-  the Mutate Search performs way better.
+  [\ldots]Suspendisse dapibus pretium eros, quis varius nisi tincidunt ac.
+Duis purus nibh, ullamcorper in ullamcorper eu, pharetra eu nunc.
+Praesent id ipsum quis est elementum ultricies elementum et odio.
+Aliquam a mi diam, at posuere lorem.
+Donec vel nisl lorem.
+Sed metus nulla, tincidunt sed ornare ac, adipiscing eu augue.
+Aenean accumsan odio ac sem rutrum feugiat.
+In hac habitasse platea dictumst.
+Mauris fringilla dui eget lectus iaculis at venenatis sem tincidunt.
+Aliquam ullamcorper nibh eget enim consectetur vel bibendum tellus molestie.
+Etiam mauris lectus, auctor vel ornare eu, accumsan nec enim.
+Vivamus blandit, est ac vehicula blandit, odio tortor semper lectus, quis luctus eros dolor a elit.
+Donec consequat tempor purus, eleifend suscipit velit rhoncus vel.
+Nam sed purus justo, id dapibus urna.
+Nullam semper mollis tempor.
 

### 2009-12-02 Paper/paper.tex
Paper update

@@
 \usepackage{epstopdf}
+\usepackage{float}
 % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-    The details of the function we used are essentially arbitrary.
+    We chose the weighting constants $w_s=2$ and $w_r=1$, which suggest that we are 
+    more concerned with the students' schedules than the rooms'.
+
+    The details of the function we used are essentially arbitrary, and we expect 
+    similar results would be attained for any other penalty-based fitness function 
+    modelling the preferences of students and requirements of rooms.
 
@@
 
-  \subsection{Options for the Genetic Algorithm}
-    \subsubsection{Scheduling Problem Instances}
+    \subsection{Scheduling Problem Instances}
       We generated 2 sets of 3 scheduling problem instances to test the algorithms above with. 
-      The package \texttt{ScheduleInstanceGenerator} in our code repository manages the generation of 
+      The package \texttt{Schedule Instance Generator} in our code repository manages the generation of 
       random problem instances with many free parameters to specify characteristics of the 
@@
 	  Toronto's Faculty of Arts and Science.
-	  The full input files for these instances are available in our repository in the 
-	  directory \emph{Instance Input Files}.
 
-    \subsubsection{Selecting Parameters for the GA}
+\section{Evaluation}
+   \subsection{Selecting Parameters for the GA}
       In order to select parameters for the genetic algorithm (that is, the population, 
@@
         the scale clearer, the colour on the surface corresponds to the fitness, where 
-        the highest fitnesses are red. The number of copies is fixed at 1, and the number 
-        of mutations (m) and crossovers (c) are shown on the horizontal axes. The total 
-        size of the population is fixed, so the number of random individuals used at each 
-        point is 60?-(m+c). This corresponds to the distance from the diagonal.
+        the highest fitnesses are red. The number of copies is fixed at 1, and the proportions 
+        of mutations and crossovers are shown on the horizontal axes.
 
-        [GRAPH]
+        \begin{figure}[H]
+  		  \centering
+  		  \includegraphics[width=8cm]{GAParametersData}
+  	      \caption{Effectiveness of the GA's proportion parameters}
+  	      \label{graph:GAParam1}
+		\end{figure}
 
@@
       \paragraph{Size Parameters}
-        Blah blah blaj. We investigated the population size and
+        We also performed tests to gauge the optimal population size and number of generations 
+        to calculate. We determined that very small values for either the population or number 
+        of generations result in very poor generated solutions. A population of around 50 is 
+        sufficient for good diversity to be retained in the population, and having a number of 
+        generations relatively large compared to the population size seemed to be effective.
 
-        [Insert fancy 3D graph here = WIN]
+        We gathered data from the small and medium problem instances, but the precise qualities 
+        provided by particular parameters will not necessarily hold in larger instances. The 
+        optimal population size and number of generations (for a fixed number of fitness 
+        evaluations) is likely to vary for larger problem instances, but it seemed clear that 
+        the number of generations should be set somewhat higher than the population.
+
+        %\begin{figure}[H]
+  		%  \centering
+  		%  \includegraphics[width=8cm]{MEDIUM-GA_Size_Data}
+  	    %  \caption{Effectiveness of the GA's population parameters}
+  	    %  \label{graph:GAParam2}
+		%\end{figure}
 
   \subsection{Comparison Between the Different Searches}
-    [\ldots]
-    \begin{figure}[h!]
+    We tested the 4 search methods on all 6 of our peoblem instances. The results for 
+    only two of the problems are shown, she small and large problems, since the medium 
+    problem exhibited characteristics very similar to the large problem.
+
+    Since all methods produced nearly indistinguishable results for the uncorrelated 
+    versions. This suggests that it may be possible to use the expected internal 
+    correlations more effectively that the methods we have proposed do.
+
+    On the graphs for the small instance, the results shown are averages over many trial 
+    runs. Individually, however, the different methods all produced similar results, in 
+    particular the ranking of the algorithms was very consistent for any particular problem 
+    instance.
+
+    \begin{figure}[H]
   	  \centering
-  	  \setlength\fboxsep{0pt}
-	  \setlength\fboxrule{0.5pt}
-	  \fbox{\includegraphics[width=8.35cm]{HUGE-TYPE2_FullAverageStats}}
-      \caption{Search comparison on a large data set}
+  	  \includegraphics[width=8cm]{SMALL-TYPE2_FullAverageStats}
+      \caption{Comparison of the different searches on a small data set}
 	\end{figure}
 
-\section{Evaluation}
-  \subsection{Success}
-    [\ldots]
-    Seeing as our fitness function gives a very high penalty for schedules that have direct
-    conflicts (either a student with two exams at the same moment, or two different exams
-    in a same room at the same moment), if such schedules exists given the input data, 
-    our algorithms will tend to very quickly reject them.
+	\begin{figure}[H]
+  	  \centering
+  	  \includegraphics[width=8cm]{HUGE-TYPE2_FullAverageStats}
+      \caption{Comparison of the different searches on a large data set}
+	\end{figure}
+
+	The mutate search algorithm performed remarkably well in all tests, finding a 
+	schedule at least somewhat better than every other method for every size problem 
+	instance.
 
@@
   [\ldots]
-  We come up with a conclusion that in this particular problem, genetic algorithm
-  is better than random search, and it is better than hill-climbing if the input 
-  is a large data set. But it may not be the best solution for this problem, since
-  the Mutate Search performs way better.
 

### 2009-12-02 Paper/paper.tex
Paper update

@@
   We examine the exam scheduling problem with soft constraints for problems of varying
-  sizes using a genetic algorithm, random search, hillclimbing, and a variant of simulated
+  sizes using a genetic algorithm (GA), random search, hillclimbing, and a variant of simulated
   annealing. This is done in the context of maximizing the fitness of a schedule, where the
@@
   algorithm is discussed in detail. We consider selecting parameters for the GA that
-  maximize the quality of the schedule found after a constant number of evaluations. The
-  GA, with a naive crossover operator, performs much better than the random search, but
+  maximize the quality of the schedule found after a constant number of fitness computations. 
+  The GA, with a naive crossover operator, performs much better than the random search, but
   simulated annealing is far superior in this search space.
@@
   poorly any schedule where she has 2 consecutive exams, and very poorly any 
-  schedule which expects her to take two exams at the same time.)
+  schedule which expects her to take two exams at the same time. A room will 
+  rate poorly a schedule that assigns two different exams to be held in it at 
+  the same moment).
 
@@
 
-      We will see that this representation is far from ideal, but there is no clear alternative.
+      We will see that this representation is far from ideal, but there is no clear 
+      alternative.
 
@@
 
-      If a different representation were chosen, such that the mutated schedules were
+      If a different representation were chosen, such that the children schedules were
       different from, but likely to have quality similar to their parents, the GA would 
@@
     \subsection{Scheduling Problem Instances}
-      We generated 2 sets of 3 scheduling problem instances to test the algorithms above with. 
-      The package \texttt{Schedule Instance Generator} in our code repository manages the generation of 
-      random problem instances with many free parameters to specify characteristics of the 
-      output instance.
+      We generated 2 sets of 3 scheduling problem instances with which to test the 
+      above algorithms. The first set of instances were generated with the parameters 
+      in the first three columns of the table below. The rooms, courses and students 
+      were given arbitrary names and the courses that each student attended were 
+      selected with uniform probability from the complete set of courses.
 
-      The first set of instances were generated with the parameters in the table below. 
-      The rooms and courses and students were given arbitrary names and the courses that each 
-      student attended were selected with uniform probability from the complete set of courses.
+      The second set was generated with a sort of correlation between the courses that each 
+      student took, rendering the dataset more true to the sort of data that would be found 
+      in real-world situations. The set of courses that each student takes was selected such 
+      that we could expect that 70\% of their courses were courses in their major. This 
+      simulates the way in which we expect student bodies to choose their courses, though 
+      the true data is likely to exhibit far more structure, due to program requirements, 
+      notorious professors, etc.
 
-      The second set was generated with a sort of correlation between the courses that each student 
-      took, rendering the dataset more true to the sort of data that would be found in real-world 
-      instances of the exam scheduling problem. The set of courses that each student takes was 
-      selected in the following way. Each student was assigned, with uniform probability, a \emph{major} 
-      and a random number of courses between 2 and 6. Each course was then selected, with probability 
-      0.7, from among the $|S|/\#Majors$ courses in their major, or, with probability 0.3, uniformly from 
-      the set of all courses. We can expect that each student takes 70\% of their courses from a small 
-      subset of the courses, representing the field that they specialize in, and 30\% from other fields. 
-      This simulates the way in which we expect student bodies to choose their courses, though the true 
-      data is likely to exhibit far more structure, due to program requirements, notorious professors, etc.
+      The 3 instances' size vary from manageable to large. The smallest could potentially be 
+      solved exactly, with methods not under consideration here, while the others have 
+      search spaces much too large for any hope of finding a perfect solution. For the 
+      largest, which simulates a UofT exam schedule, a single evaluation of the fitness 
+      function takes 0.2 seconds on on a typical desktop computer. This makes it difficult 
+      to thoroughly investigate this instance.
 
-      The 3 sizes vary from manageable to large. The smallest could potentially be solved exactly, 
-      with methods not under consideration here, while the others have search spaces much too large 
-      for any hope of an exact solution. For the largest, which simulates a UofT exam schedule, a 
-      single evaluation of the fitness function took 0.2 seconds on the fastest machine owned by any 
-      of the authors.
-
-      The following table characterized the problem instances that we used 
-      (with \textsc{SSS} being an approximation of the search space size):      
+      The following table characterizes the problem instances that we used ($\#$Maj is the 
+      number of different majors used and \textsc{SSS} is an approximation of the search space size):    
       \definecolor{light-gray}{gray}{0.87}
@@
 
-        In the graph below, each data point represents the average of 5 trials. To make 
-        the scale clearer, the colour on the surface corresponds to the fitness, where 
-        the highest fitnesses are red. The number of copies is fixed at 1, and the proportions 
-        of mutations and crossovers are shown on the horizontal axes.
+        Our results can be seen on figure 1, where each data point of the graph 
+        represents the average of 5 trials. To make the scale clearer, the colour 
+        on the surface corresponds to the fitness, where the highest fitnesses 
+        are red. The number of copies is fixed at 1, and the proportions of 
+        mutations and crossovers are shown on the horizontal axes.
 
-        \begin{figure}[H]
+        \begin{figure}[h]
   		  \centering
@@
 
-        The graph above indicates that for a fixed number of evaluations, best results are 
+        The graph indicates that for a fixed number of evaluations, best results are 
         achieved by selecting parameters that have the majority of individuals in the next 
@@
 
-        %\begin{figure}[H]
-  		%  \centering
-  		%  \includegraphics[width=8cm]{MEDIUM-GA_Size_Data}
-  	    %  \caption{Effectiveness of the GA's population parameters}
-  	    %  \label{graph:GAParam2}
-		%\end{figure}
+  \subsection{Comparison Between the Different Searches}
+    We tested the 4 search methods on all 6 of our problem instances. The results 
+    for two of the problems, the small and large problems, are shown on figure 2 
+    and 3. The medium problem exhibited characteristics very similar to the large problem.
 
-  \subsection{Comparison Between the Different Searches}
-    We tested the 4 search methods on all 6 of our peoblem instances. The results for 
-    only two of the problems are shown, she small and large problems, since the medium 
-    problem exhibited characteristics very similar to the large problem.
-
-    Since all methods produced nearly indistinguishable results for the uncorrelated 
+    All methods produced nearly indistinguishable results for the uncorrelated 
     versions. This suggests that it may be possible to use the expected internal 
@@
 
-    \begin{figure}[H]
+    \begin{figure}[h]
   	  \centering
@@
 
-	\begin{figure}[H]
+	\begin{figure}[h]
   	  \centering
@@
 
+	Random search provides a benchmark by which to measure the other 3 techniques. 
+	As desired, they all usually perform significantly better.
+
+	Hillclimbing works very well for the small instance, but miserably on the large 
+	problem. This is because of the vastly increased number of neighbouring schedules 
+	in the large problem. Hillclimbing is able to examine a much smaller portion of 
+	search space and wastes its evaluations on too many points that are too similar 
+	to each other. A variant that tries some fixed number of schedules in the 
+	neighbourhood may perform better.
+
 	The mutate search algorithm performed remarkably well in all tests, finding a 
@@
 
+	The GA is a bit of a disappointment. It works much better than random, meaning 
+	that the naive crossover method used does have some value, and it scales well 
+	for different sized problem instances, which is an important characteristic. 
+	For the large instance, the attained fitness of around 0.4 would be the fitness 
+	of a schedule for which every student and room assign 1.4 penalty points to the 
+	schedule. This is reasonable, and would perhaps be very exciting if mutate search 
+	wasn't so much better than GA for this problem.
+
 \section{Conclusion}
-  [\ldots]
+  \subsection{Fitness Function Modifications}
+    The fitness function can be modified by any monotone function to obtain identical 
+    results for the 3 non-GA methods, which only need to choose the better fitness 
+    and do not use the fitness values in any important way. The GA, however, uses the 
+    fitness values to decide the probability of schedules being selected and modified 
+    for subsequent generations. Modifying the fitness function in a variety of ways 
+    (perhaps calculating the total penalty that a schedule gets instead averaging over 
+    the values given by each student and room) might improve the quality of the GA 
+    approach to this problem. The fitness function used here could be modified in many 
+    ways, and investigating this would be very interesting.
 

### 2009-12-02 Paper/paper.tex
Paper update

@@
 \section{Conclusion}
-  \subsection{Fitness Function Modifications}
-    The fitness function can be modified by any monotone function to obtain identical 
+  Any monotone function can be applied to the fitness values to obtain identical 
   results for the 3 non-GA methods, which only need to choose the better fitness 
-    and do not use the fitness values in any important way. The GA, however, uses the 
-    fitness values to decide the probability of schedules being selected and modified 
-    for subsequent generations. Modifying the fitness function in a variety of ways 
-    (perhaps calculating the total penalty that a schedule gets instead averaging over 
-    the values given by each student and room) might improve the quality of the GA 
-    approach to this problem. The fitness function used here could be modified in many 
-    ways, and investigating this would be very interesting.
+  and do not use the fitness values in any important way. The GA, however, uses 
+  the fitness values to decide the probability of schedules being selected and 
+  modified for subsequent generations. It might be possible to reduce or increase 
+  the probability of poor functions being selected for mutation or crossover.
+
+  Modifying the fitness function in a variety of ways (perhaps calculating the 
+  total penalty that a schedule gets instead averaging over the values given by 
+  each student and room) might improve the  quality of the GA approach to this 
+  problem. The fitness function used here could  be modified in many ways, and 
+  investigating this would be very interesting.
+
+  Unfortunately, without further investigation and modification, the GA given 
+  should not be used. Perhaps a novel form for representing schedules could make 
+  a GA approach for this problem perform better than mutate search. We recommend 
+  Mutate search for solving problems of this type. A typical PC can generate a 
+  relatively good schedule for a large exam scheduling problem overnight using the
+  mutate search method. Mutate search is very easy to implement, making it even 
+  more attractive.
+
+  Mutate search can be modified to change the rate of mutation and number of attempts 
+  at each step, which suggests that much better methods can be developed based on the 
+  it. This could be an interesting direction for further research.
 

### 2009-12-02 Paper/paper.tex
Paper update

@@
 \usepackage{float}
+\usepackage{natbib}
 % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-  The total number of different possible schedules in any problem instance is
-  $|C|^{|R|dt}$ which is typically far too large for brute force search.
-  For example, one instance that we examine (scheduling the fall semester
-  exams at the University of Toronto in 2009) involves 603 courses, 7 days,
-  8 times, 43 rooms, and 21945 students. This instance admits approximately 
-  $10^{6695}$ different possible schedules.
-
   Typically, with relatively loose constraints on the number of rooms and 
@@
   \subsection{Differences From Natural Exam Scheduling Problems}
-    The problem we have constructed here is a compromise between the most
-    mathematically pure problem with the same sort of properties (which might ignore the
-    rooms and related constraints completely) and the problem that must be solved in the
-    real world when scheduling thousands of students, which has some important differences
-    from this problem. For example, we don't deal with constraints such as room capacity
-    or exams with varying length.
+    The problem we have constructed here is a compromise between more mathematically
+    pure problems with the same sort of properties (which might ignore the rooms' 
+    constraints completely) and the problem that must be solved in the real world 
+    when scheduling thousands of students, which has some important differences from 
+    this problem. For example, we don't deal with constraints like room capacity or 
+    exams with varying length.
 
@@
     to generate problem instance files for testing our algorithms. We used Mathematica 
-    to plot all of our graphs.  Additionally, we use Google code's source code
-    repository\footnote{\url{http://code.google.com/p/csc384-genetic-algorithms-project/}} 
-    and Google docs to collaborate on this report.
+    to plot all of our graphs.  Additionally, we use Google code's subversion
+    repository\footnote{\url{http://code.google.com/p/csc384-genetic-algorithms-project/}},
+    Google docs and Google Wave to collaborate on this report.
 
@@
     In order to assign each schedule a fitness, we need to define the functions
-    $q_{student}$ and $q_{room}$. They need to capture the idea of determining how
-    much each student or room likes their given timetable. We do this by calculating 
-    the penalty incurred by a room, which is a measure of it having properties that 
-    are unpleasant to the student or room, and then using the formula
+    $q_{student}$ and $q_{room}$. They need to capture the idea of determining 
+    how much each student or room likes their given timetable. We do this by giving 
+    penalties to schedules when they have properties that are disliked by the 
+    student or room, and then using the formula
 
@@
 
-    This function will be in the desired range or [0,1].
+    This function will be in the desired range of [0,1].
 
-    For students, we accrue a penalty of 50 for every time that is repeated in a 
-    timetable, 5 when there are exams in 2 consecutive times, 3 when 2 exams happen 
-    on the same day, and 0.5 when the student has exams on consecutive days. This 
-    accords reasonably well with real students' hopes for their exam timetables.
-
-    Rooms have the same penalty of 50 when it is used by two exams at the same time, 
-    5 if it hosts exams in consecutive time slots (reasoning that the invigilators 
-    need time to let the students out and in and prepare for the next exam) and a 
-    penalty of 0.5 when a room is empty for an entire day between uses (reasoning 
-    that the room requires some amount of modifications to be used for exams, and 
-    it is difficult to use it for other purposes if it will still be used for exams 
-    soon.)
+    We assign 50 point penalties for inconsistent schedules, 3-5 points when 
+    students are assigned multiple exams on the same day, and 0.5 when students 
+    don't have a rest day between exams. Rooms have the same stiff 50 point penalty 
+    for inconsistencies, 5 for hosting exams in consecutive time slots, and 0.5 for 
+    any days without exams between the days it is first used and last used.
 
@@
     genetic algorithm package, in Java, that can interface with data types designed 
-    by others, so long as they implement a few necessary methods. The package we wrote 
-    is more than adequate for investigating the scheduling problem under consideration here.
+    by others, so long as they implement a few necessary methods. The package we 
+    wrote is more than adequate for investigating the scheduling problem under 
+    consideration here.
 
@@
       the schedules with high quality to be selected more frequently and make more 
-      ``offspring''.
+      ``offspring".
 
-      The free parameters in our GA are the size of the population, number of
-      generations to simulate, and proportions of new generations to be made by the copy, 
-      random, mutation, and crossover methods.
+      The free parameters in our GA are the size of the population, number of generations 
+      to simulate, and proportions of new generations to be made by the copy, random, 
+      mutation, and crossover methods.
 
@@
 
+    \subsubsection{Parallel Computations}
+      Of the 4 methods, only mutate search is not trivially parallelizable. A plausible 
+      parallel version of mutate search would generate as many mutated schedules at 
+      every step as processors are available and select the fittest result for the next 
+      set of mutations. We expect that this would increase the rate at which the best 
+      found fitness increases.
+
   \subsection{Scheduling Problem Instances}
@@
 
-      The second set was generated with a sort of correlation between the courses that each 
-      student took, rendering the dataset more true to the sort of data that would be found 
-      in real-world situations. The set of courses that each student takes was selected such 
-      that we could expect that 70\% of their courses were courses in their major. This 
-      simulates the way in which we expect student bodies to choose their courses, though 
-      the true data is likely to exhibit far more structure, due to program requirements, 
-      notorious professors, etc.
+    The second set was generated with a correlation between the courses that each 
+    student took, rendering the data-set more true to the sort of data that would 
+    be found in real-world situations. The set of courses that each student takes 
+    was selected so that around 70\% of their courses were in their (randomly selected) 
+    major. This simulates the way in which we expect student bodies to choose their 
+    courses, though the true data is likely to exhibit far more structure, due to 
+    program requirements, notorious professors, etc.
 
-      The 3 instances' size vary from manageable to large. The smallest could potentially be 
-      solved exactly, with methods not under consideration here, while the others have 
-      search spaces much too large for any hope of finding a perfect solution. For the 
-      largest, which simulates a UofT exam schedule, a single evaluation of the fitness 
-      function takes 0.2 seconds on on a typical desktop computer. This makes it difficult 
-      to thoroughly investigate this instance.
+    The sized of our 3 instances' vary from manageable to large. The total number 
+    of different possible schedules in any problem instance is $|C|^{|R|dt}$ which 
+    is far too large to conduct a brute force search. The smallest could potentially 
+    be solved exactly, with methods not under consideration here, while the others 
+    have search spaces much too large for any hope of finding a perfect solution. 
+    For the largest, which simulates a UofT exam schedule, a single evaluation of 
+    the fitness function takes 0.2 seconds on on a typical desktop computer. This 
+    makes it difficult to thoroughly investigate this instance.
 
@@
 
-	  The large data set was created with parameters inspired from the University of 
-	  Toronto's Faculty of Arts and Science.
+	The large data set was created with parameters corresponding to the fall 2009 
+	University of Toronto exam schedule, which is the size we expect for a large 
+	real-world problem instance.
 
@@
         We undertook a search of the space of parameters for forming new generations. 
-
-        Our results can be seen on figure 1, where each data point of the graph 
-        represents the average of 5 trials. To make the scale clearer, the colour 
-        on the surface corresponds to the fitness, where the highest fitnesses 
-        are red. The number of copies is fixed at 1, and the proportions of 
-        mutations and crossovers are shown on the horizontal axes.
+        Our results for different combinations of parameters can be seen in figure 1. 
+        Each point has been averaged over 5 trials. The colour on the plotted surface 
+        corresponds to the fitness, where the highest fitnesses are red. The number of 
+        copies is fixed at 1, and the proportion of mutations and crossovers are shown 
+        on the horizontal axes.
 
@@
         preference.
-
-      \paragraph{Size Parameters}
-        We also performed tests to gauge the optimal population size and number of generations 
-        to calculate. We determined that very small values for either the population or number 
-        of generations result in very poor generated solutions. A population of around 50 is 
-        sufficient for good diversity to be retained in the population, and having a number of 
-        generations relatively large compared to the population size seemed to be effective.
-
-        We gathered data from the small and medium problem instances, but the precise qualities 
-        provided by particular parameters will not necessarily hold in larger instances. The 
-        optimal population size and number of generations (for a fixed number of fitness 
-        evaluations) is likely to vary for larger problem instances, but it seemed clear that 
-        the number of generations should be set somewhat higher than the population.
 
@@
   the probability of poor functions being selected for mutation or crossover.
+  \citeauthor*{pdf:solvingexam} \citeyearpar{pdf:solvingexam} for example, got 
+  favourable results by squaring the fitness.
 

### 2009-12-02 trunk/Paper/paper.tex
Updated paper with some new graphs
Paper update

--- 2009-11-30 trunk/Paper/paper.tex
@@
 \usepackage{url}
+\usepackage[table]{xcolor}
+\usepackage{graphicx}
+\usepackage{epstopdf}
+\usepackage{float}
+\usepackage{natbib}
 % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   which we can express like this:
-
   \[ C \rightarrow \left\{(r,\ a,\ b)\ |\ r\in R,\ a\in \{1..d\},\ b\in \{1..t\}\right\} \]
-
-  The total number of different possible schedules in any problem instance is
-  $|C|^{|R|dt}$ which is typically far too large for brute force search.
-  For example, one instance that we examine (scheduling the fall semester
-  exams at the University of Toronto in 2009) involves 603 courses, 7 days,
-  8 times, 43 rooms, and 21945 students. This instance admits approximately 
-  $10^{6695}$ different possible schedules.
-
   Typically, with relatively loose constraints on the number of rooms and 
@@
   $K$ may be represented in this way:
-
   \[ TT_s(K) = \left\{(a, b)\ |\ \exists r \in R, \exists c \in L_s, K(c)=(r, a, b) \right\} \]
-
   which may, in general, be a multiset.
@@
   Similarly, the timetable for a room $r \in R$ is
-
   \[ TT_r(K)=\left\{(a, b)\ |\ \exists c\in C,\ K(c)=(r, a, b)\right\} \]
@@
   one for students and one for rooms. These are meant to capture how much they
-  "like" the schedule under consideration. (For instance, a student will rate
+  ``like" the schedule under consideration. (For instance, a student will rate
   poorly any schedule where she has 2 consecutive exams, and very poorly any 
-  schedule which expects her to take two exams at the same time.)
-
+  schedule which expects her to take two exams at the same time. A room will 
+  rate poorly a schedule that assigns two different exams to be held in it at 
+  the same moment).
   \[ Q_s(K)=q_{student}\left(TT_s(K)\right):\textit{ Schedules}\rightarrow [0,\ 1] \]
   \[ Q_r(K)=q_{room}\left(TT_r(K)\right):\textit{ Schedules}\rightarrow [0,\ 1] \]
-
   We will consider later how to define these functions in a reasonable way.
@@
   The quality of a schedule, then, is given by
-
   \[ Q(K)= \frac{\left(w_s\left(\sum_{s \in S}{\frac{Q_s(K)}{|S|}}\right) +
      w_r\left(\sum_{r\in R}{\frac{Q_r(K)}{|R|}} \right)\right)}{(w_s+w_r)} \]
-
   where $w_s$ and $w_r$ are the weightings we can use to indicate that we 
@@
 \section{Approach and Implementation}
-  \subsection{Differences From Natural Exam Scheduling Problems}
-    The problem we have constructed here is a compromise between the most
-    mathematically pure problem with the same sort of properties (which might ignore the
-    rooms and related constraints completely) and the problem that must be solved in the
-    real world when scheduling thousands of students, which has some important differences
-    from this problem.
+  \subsection{Differences From Natural Exam Scheduling}
+    The problem we have constructed here is a compromise between more mathematically
+    pure problems with the same sort of properties (which might ignore the rooms' 
+    constraints completely) and the problem that must be solved in the real world 
+    when scheduling thousands of students, which has some important differences from 
+    this problem. For example, we don't deal with constraints like room capacity or 
+    exams with varying length.
 
   \subsection{Tools used}
-    The program is implemented in Java, to generate possible schedules and to calculate
-    their fitness. A Mathematica program is also written to plot graphs, given the data
-    generated by the Java program. Additionally, we use Google code 
-    subversion\footnote{\url{http://code.google.com/p/csc384-genetic-algorithms-project/}}
-    and Google doc to share the work.
+    Our application was entirely written in Java. We also wrote another Java application 
+    to generate problem instance files for testing our algorithms. We used Mathematica 
+    to plot all of our graphs.  Additionally, we use Google code's subversion
+    repository\footnote{\url{http://code.google.com/p/csc384-genetic-algorithms-project/}},
+    Google docs and Google Wave to collaborate on this report.
 
@@
     In order to assign each schedule a fitness, we need to define the functions
-    $q_{student}$ and $q_{room}$. They need to capture the idea of determining how
-    much each student or room likes their given timetable. We do this by calculating 
-    the penalty incurred by a room, which is a measure of it having properties that 
-    are unpleasant to the student or room, and then using the formula
+    $q_{student}$ and $q_{room}$. They need to capture the idea of determining 
+    how much each student or room likes their given timetable. We do this by giving 
+    penalties to schedules when they have properties that are disliked by the 
+    student or room, and then using the formula
+    \[ q(TT)=\frac{1}{1+penalty_{TT}} \]
+    This function will be in the desired range of [0,1].
 
-    \[ q(TT)=\frac{1}{1+penalty_{TT}} \]
+    We assign 50 point penalties for inconsistent schedules, 3-5 points when 
+    students are assigned multiple exams on the same day, and 0.5 when students 
+    don't have a rest day between exams. Rooms have the same stiff 50 point penalty 
+    for inconsistencies, 5 for hosting exams in consecutive time slots, and 0.5 for 
+    any days without exams between the days it is first used and last used.
 
-    This function will be in the desired range or [0,1].
+    We chose the weighting constants $w_s=2$ and $w_r=1$, which suggest that we are 
+    more concerned with the students' schedules than the rooms'.
 
-    For students, we accrue a penalty of 50 for every time that is repeated in a 
-    timetable, 5 when there are exams in 2 consecutive times, 3 when 2 exams happen 
-    on the same day, and 0.5 when the student has exams on consecutive days. This 
-    accords reasonably well with real students' hopes for their exam timetables.
-
-    Rooms have the same penalty of 50 when it is used by two exams at the same time, 
-    5 if it hosts exams in consecutive time slots (reasoning that the invigilators 
-    need time to let the students out and in and prepare for the next exam) and a 
-    penalty of 0.5 when a room is empty for an entire day between uses (reasoning 
-    that the room requires some amount of modifications to be used for exams, and 
-    it is difficult to use it for other purposes if it will still be used for exams 
-    soon.)
-
-    The details of the function we used are essentially arbitrary.
+    The details of the function we used are essentially arbitrary, and we expect 
+    similar results would be attained for any other penalty-based fitness function 
+    modelling the preferences of students and requirements of rooms.
 
@@
     Although there are many readily available open source genetic algorithm software 
-    packages available (see JGAP \cite{website:JGAP}, Jenes\cite{website:jenes}) we 
-    decided to write a simple, yet general, genetic algorithm package, in Java, that
-    can interface with data types designed by others, so long as they implement a few
-    necessary methods. The package we wrote is more than adequate for investigating
-    the scheduling problem under consideration here.
+    packages available (see JGAP, Jenes) we decided to write a simple, yet general, 
+    genetic algorithm package, in Java, that can interface with data types designed 
+    by others, so long as they implement a few necessary methods. The package we 
+    wrote is more than adequate for investigating the scheduling problem under 
+    consideration here.
 
@@
       schedules are selected with a probability proportional to their fitness, allowing 
-      the schedules with high quality tobe selected more frequently and make more 
-      "offspring".
+      the schedules with high quality to be selected more frequently and make more 
+      ``offspring".
 
-      The free parameters in our GA are the size of the population, number of
-      generations to simulate, and proportions of new generations toi be made by the copy, 
-      random, mutation, and crossover methods.
+      The free parameters in our GA are the size of the population, number of generations 
+      to simulate, and proportions of new generations to be made by the copy, random, 
+      mutation, and crossover methods.
 
@@
       (corresponding to a pair or integers representing the day and time) of each course's
-      exam.
+      exam. We will see that this representation is far from ideal, but there is no clear 
+      alternative.
 
@@
       The neighbouring schedules are those where only one course meets in either a different
-      room, on a different day, or at a different time. This leads to a total of
-      $ |C|(d+t+|R|-3) $ schedules in the neighbourhood of any given schedule.
-
+      room, on a different day, or at a different time. 
+%      This leads to a total of $ |C|(d+t+|R|-3) $ schedules in the neighbourhood of any given schedule.
       It can be seen that this method is completely deterministic given a starting schedule,
-      and will always return a local maximum
+      and will always return a local maximum.
 
@@
       (by using the same mutate operator made for the genetic algorithm) with 1\% of the
-      data randomized and chooses which of the schedules is better. This operation is
+      data randomized and keeps the better schedule. This operation is
       repeated until it has performed as many evaluations as it is allowed to.
@@
 
-  \subsection{Options for the Genetic Algorithm}
-    \subsubsection{Scheduling Problem Instances}
-      We generated 2 sets of 3 scheduling problem instances to test the algorithms above with. 
-      The package ScheduleInstanceGenerator in our code repository manages the generation of 
-      random problem instances with many free parameters to specify characteristics of the 
-      output instance.
+    \subsubsection{Parallel Computations}
+      Of the 4 methods, only mutate search is not trivially parallelizable. A plausible 
+      parallel version of mutate search would generate as many mutated schedules at 
+      every step as processors are available and select the fittest result for the next 
+      set of mutations. We expect that this would increase the rate at which the best 
+      found fitness increases.
 
-      The following table characterized the problem instances that we used:
+  \subsection{Scheduling Problem Instances}
+    We generated 2 sets of 3 scheduling problem instances with which to test the 
+    above algorithms. The first set of instances were generated with the parameters 
+    in the first three columns of the table below. The rooms, courses and students 
+    were given arbitrary names and the courses that each student attended were 
+    selected with uniform probability from the complete set of courses.
 
-      \begin{tabular}{|c|c|c|c|c|c|c|}
-      \hline  & Tiny & Medium & Large & Tiny /w Cor. & Med /w Cor. & Large /w Cor. \\ 
-      \hline d & 7 & 10 & 7 & 7 & 10 & 7 \\ 
-      \hline t & 5 & 8 & 8 & 5 & 8 & 8 \\ 
-      \hline $|C|$ & 20 & 200 & 603 & 20 & 200 & 600 \\ 
-      \hline $|R|$ & 4 & 10 & 43 & 4 & 10 & 43 \\ 
-      \hline $|S|$ & 50 & 300 & 21945 & 50 & 300 & 21945 \\ 
-      \hline $|L_s|$ & 1-5 & 1-6 & 1-5 & 2-6 & 2-6 & 2-6 \\ 
-      \hline \# Course Types & N/A & N/A & N/A & 4 & 10 & 20 \\ 
+    The second set was generated with a correlation between the courses that each 
+    student took, rendering the data-set more true to the sort of data that would 
+    be found in real-world situations. The set of courses that each student takes 
+    was selected so that around 70\% of their courses were in their (randomly selected) 
+    major. This simulates the way in which we expect student bodies to choose their 
+    courses, though the true data is likely to exhibit far more structure, due to 
+    program requirements, notorious professors, etc.
+
+    The sized of our 3 instances' vary from manageable to large. The total number 
+    of different possible schedules in any problem instance is $|C|^{|R|dt}$ which 
+    is far too large to conduct a brute force search. The smallest could potentially 
+    be solved exactly, with methods not under consideration here, while the others 
+    have search spaces much too large for any hope of finding a perfect solution. 
+    For the largest, which simulates a UofT exam schedule, a single evaluation of 
+    the fitness function takes 0.2 seconds on on a typical desktop computer. This 
+    makes it difficult to thoroughly investigate this instance.
+
+    The large data set was created with parameters corresponding to the fall 2009 
+	University of Toronto exam schedule, which is the size we expect for a large 
+	real-world problem instance.
+
+    \definecolor{light-gray}{gray}{0.87}
+    \rowcolors{1}{light-gray}{white}
+    \begin{table}[h]
+    \begin{small}
+    \begin{tabular}{*{7}{c}}
+    \hline 
+ 	  	& S 	& M 	& L 	& $S_{Cor}$ & $M_{Cor}$ & $L_{Cor}$ \\ 
+    \hline \hline
+     d 		& 7 	& 10	& 7 	& 7 	& 10 	& 7 	\\ 
+     t 		& 5		& 8 	& 8 	& 5 	& 8 	& 8 	\\ 
+     $|C|$	& 20 	& 200 	& 603 	& 20 	& 200 	& 600 	\\ 
+     $|R|$	& 4 	& 10 	& 43 	& 4 	& 10 	& 43	\\ 
+     $|S|$	& 50 	& 300 	& 21945 & 50 	& 300 	& 21945	\\ 
+     $|L_s|$	& 1-5 	& 1-6 	& 1-5 	& 2-6 	& 2-6 	& 2-6 	\\ 
+     \begin{tiny}$\#$Maj\end{tiny}	& N/A 	& N/A 	& N/A 	& 4 	& 10 	& 20 	\\ 
+     \textsc{SSS}	& $10^{182}$ & $10^{1840}$ & $10^{6695}$ & $10^{182}$ & $10^{1840}$ & $10^{6689}$ \\
     \hline  
     \end{tabular} 
-
-	  The full input files for these instances are available in our repository in the 
-	  directory "Instance Input Files".
-
-    \subsubsection{Selecting Parameters for the GA}
-      [\ldots]
-  \subsection{Comparison Between the Different Searches}
-    [\ldots]
+	\end{small}
+	\caption{The six data sets used in our project. $\#$Maj is the number of different 
+	         majors used and \textsc{SSS} is an approximation of the search space size.}
+	\end{table}
 
 \section{Evaluation}
-  \subsection{Success}
-    [\ldots]
-    Seeing as our fitness function gives a very high penalty for schedules that have direct
-    conflicts (either a student with two exams at the same moment, or two different exams
-    in a same room at the same moment), if such a schedule exists given the input data, our
-    algorithms will very quickly tend to reject all such schedules.
+   \subsection{Selecting Parameters for the GA}
+      In order to select parameters for the genetic algorithm (that is, the population, 
+      number of generations to simulate, proportions of copies, randoms, mutations, and 
+      crossovers for the next generation) we performed tests using the small uncorrelated 
+      problem instance.
+
+      \paragraph{Generation Parameters}
+        We undertook a search of the space of parameters for forming new generations. 
+        Our results for different combinations of parameters can be seen in figure 1. 
+        Each point has been averaged over 5 trials. The colour on the plotted surface 
+        corresponds to the fitness, where the highest fitnesses are red. The number of 
+        copies is fixed at 1, and the proportion of mutations and crossovers are shown 
+        on the horizontal axes.
+
+        \begin{figure}[h]
+  		  \centering
+  		  \includegraphics[width=8cm]{GAParametersData}
+  	      \caption{Effectiveness of the GA's proportion parameters}
+  	      \label{graph:GAParam1}
+		\end{figure}
+
+        The graph indicates that for a fixed number of evaluations, best results are 
+        achieved by selecting parameters that have the majority of individuals in the next 
+        generation be generated by crossovers of the most fit schedules in the previous 
+        generation. In order to achieve the very highest fitness, a few randoms or mutations 
+        should also be incorporated.
+
+        This test was done for only the smallest problem instance, and is infeasible for the 
+        larger instances. We chose generation proportion parameters to approximate the optimal 
+        values seen here, under the assumption that the larger problems exhibited this sort of 
+        preference.
+
+  \subsection{Comparison Between the Different Searches}
+    We tested the 4 search methods on all 6 of our problem instances. The results 
+    for two of the problems, the small and large problems, are shown in figure 2 
+    and 3. The medium problem exhibited characteristics very similar to the large problem.
+
+    All methods produced nearly indistinguishable results for the uncorrelated 
+    versions. This suggests that it may be possible to use the expected internal 
+    correlations more effectively that the methods we have proposed do.
+
+    On the graphs for the small instance, the results shown are averages over many trial 
+    runs. Individually, however, the different methods all produced similar results, in 
+    particular the ranking of the algorithms was very consistent for any particular problem 
+    instance.
+
+    \begin{figure}[h]
+  	  \centering
+  	  \includegraphics[width=8cm]{SMALL-TYPE2_FullAverageStats}
+      \caption{Comparing the four searches on a small data set}
+	\end{figure}
+
+	\begin{figure}[h]
+  	  \centering
+  	  \includegraphics[width=8cm]{HUGE-TYPE2_FullAverageStats}
+      \caption{Comparing the four searches on a large data set}
+	\end{figure}
+
+	Random search provides a benchmark by which to measure the other 3 techniques. 
+	As desired, they all usually perform significantly better.
+
+	Hillclimbing works very well for the small instance, but miserably on the large 
+	problem. This is because of the vastly increased number of neighbouring schedules 
+	in the large problem. Hillclimbing is able to examine a much smaller portion of 
+	search space and wastes its evaluations on too many points that are too similar 
+	to each other. A variant that tries some fixed number of schedules in the 
+	neighbourhood may perform better.
+
+	The mutate search algorithm performed remarkably well in all tests, finding a 
+	schedule at least somewhat better than every other method for every size problem 
+	instance.
+
+	The GA is a bit of a disappointment. It works much better than random, meaning 
+	that the naive crossover method used does have some value, and it scales well 
+	for different sized problem instances, which is an important characteristic. 
+	For the large instance, the attained fitness of around 0.4 would be the fitness 
+	of a schedule for which every student and room assign 1.4 penalty points to the 
+	schedule. This is reasonable, and would perhaps be very exciting if mutate search 
+	wasn't so much better than GA for this problem.
 
 \section{Conclusion}
-  We come up with a conclusion that in this particular problem, genetic algorithm
-  is better than random search, and it is better than hill-climbing if the input 
-  is a large data set. But it may not be the best solution for this problem, since
-  the Mutate Search performs way better.
+  Any monotone function can be applied to the fitness values to obtain identical 
+  results for the 3 non-GA methods, which only need to choose the better fitness 
+  and do not use the fitness values in any important way. The GA, however, uses 
+  the fitness values to decide the probability of schedules being selected and 
+  modified for subsequent generations. It might be possible to reduce or increase 
+  the probability of poor functions being selected for mutation or crossover.
+  \citeauthor*{pdf:solvingexam} \citeyearpar{pdf:solvingexam} for example, got 
+  favourable results by squaring the fitness.
 
+  Modifying the fitness function in a variety of ways (perhaps calculating the 
+  total penalty that a schedule gets instead averaging over the values given by 
+  each student and room) might improve the  quality of the GA approach to this 
+  problem. The fitness function used here could  be modified in many ways, and 
+  investigating this would be very interesting.
+
+  Unfortunately, without further investigation and modification, the GA given 
+  should not be used. Perhaps a novel form for representing schedules could make 
+  a GA approach for this problem perform better than mutate search. We recommend 
+  Mutate search for solving problems of this type. A typical PC can generate a 
+  relatively good schedule for a large exam scheduling problem overnight using the
+  mutate search method. Mutate search is very easy to implement, making it even 
+  more attractive.
+
+  Mutate search can be modified to change the rate of mutation and number of attempts 
+  at each step, which suggests that much better methods can be developed based on the 
+  it. This could be an interesting direction for further research.
+
+\nocite{*}
 \bibliographystyle{aaai}

### 2009-12-02 Paper/paper.tex
Paper update

@@
   which we can express like this:
-
   \[ C \rightarrow \left\{(r,\ a,\ b)\ |\ r\in R,\ a\in \{1..d\},\ b\in \{1..t\}\right\} \]
-
   Typically, with relatively loose constraints on the number of rooms and 
@@
   one for students and one for rooms. These are meant to capture how much they
-  "like" the schedule under consideration. (For instance, a student will rate
+  ``like" the schedule under consideration. (For instance, a student will rate
   poorly any schedule where she has 2 consecutive exams, and very poorly any 
@@
   the same moment).
-
   \[ Q_s(K)=q_{student}\left(TT_s(K)\right):\textit{ Schedules}\rightarrow [0,\ 1] \]
   \[ Q_r(K)=q_{room}\left(TT_r(K)\right):\textit{ Schedules}\rightarrow [0,\ 1] \]
-
   We will consider later how to define these functions in a reasonable way.
@@
 \section{Approach and Implementation}
-  \subsection{Differences From Natural Exam Scheduling Problems}
+  \subsection{Differences From Natural Exam Scheduling}
     The problem we have constructed here is a compromise between more mathematically
@@
     student or room, and then using the formula
-
     \[ q(TT)=\frac{1}{1+penalty_{TT}} \]
-
     This function will be in the desired range of [0,1].
@@
       (corresponding to a pair or integers representing the day and time) of each course's
-      exam.
-
-      We will see that this representation is far from ideal, but there is no clear 
+      exam. We will see that this representation is far from ideal, but there is no clear 
       alternative.
@@
       The neighbouring schedules are those where only one course meets in either a different
-      room, on a different day, or at a different time. This leads to a total of
-      $ |C|(d+t+|R|-3) $ schedules in the neighbourhood of any given schedule.
+      room, on a different day, or at a different time. 
+      This leads to a total of $ |C|(d+t+|R|-3) $ schedules in the neighbourhood of any given schedule.
 
       It can be seen that this method is completely deterministic given a starting schedule,
-      and will always return a local maximum
+      and will always return a local maximum.
 
@@
       remaining. Not being the focus of our investigation, little effort was made to 
-      optimize the parameters. 1\% does seem to give better performance than other nearby 
-      values, though we did not collect data to substantiate this. Finding ideal parameters 
-      for simulated annealing on this problem would make for a project-sized inquiry in itself.
+      optimize the parameters. 
+      %1\% does seem to give better performance than other nearby 
+%      values, though we did not collect data to substantiate this. Finding ideal parameters 
+ %     for simulated annealing on this problem would make for a project-sized inquiry in itself.
 
@@
 
-    The following table characterizes the problem instances that we used ($\#$Maj is the 
-    number of different majors used and \textsc{SSS} is an approximation of the search space size):    
+    The large problem instance was created with parameters corresponding to the fall 2009 
+	University of Toronto exam schedule, which is the size we expect for a large 
+	real-world problem instance.
+
     \definecolor{light-gray}{gray}{0.87}
     \rowcolors{1}{light-gray}{white}
-    \begin{center}
+    \begin{table}[h]
     \begin{small}
@@
 	\end{small}
-	\end{center}
-
-	The large data set was created with parameters corresponding to the fall 2009 
-	University of Toronto exam schedule, which is the size we expect for a large 
-	real-world problem instance.
+	\caption{The six problem instances used in our project. $\#$Maj is the number of different 
+	         majors used and \textsc{SSS} is an approximation of the search space size.}
+	\end{table}
 
@@
     We tested the 4 search methods on all 6 of our problem instances. The results 
-    for two of the problems, the small and large problems, are shown on figure 2 
+    for two of the problems, the small and large problems, are shown in figure 2 
     and 3. The medium problem exhibited characteristics very similar to the large problem.
@@
   	  \includegraphics[width=8cm]{SMALL-TYPE2_FullAverageStats}
-      \caption{Comparison of the different searches on a small data set}
+      \caption{Search methods on small problem instance}
 	\end{figure}
@@
   	  \includegraphics[width=8cm]{HUGE-TYPE2_FullAverageStats}
-      \caption{Comparison of the different searches on a large data set}
+      \caption{Search methods on large problem instance}
 	\end{figure}

### 2009-12-02 trunk/Paper/paper.tex
Paper update

@@
       room, on a different day, or at a different time. 
-%      This leads to a total of $ |C|(d+t+|R|-3) $ schedules in the neighbourhood of any given schedule.
+      This leads to a total of $ |C|(d+t+|R|-3) $ schedules in the neighbourhood of any given schedule.
+
       It can be seen that this method is completely deterministic given a starting schedule,
@@
 
-    The sized of our 3 instances' vary from manageable to large. The total number 
-    of different possible schedules in any problem instance is $|C|^{|R|dt}$ which 
-    is far too large to conduct a brute force search. The smallest could potentially 
-    be solved exactly, with methods not under consideration here, while the others 
-    have search spaces much too large for any hope of finding a perfect solution. 
-    For the largest, which simulates a UofT exam schedule, a single evaluation of 
-    the fitness function takes 0.2 seconds on on a typical desktop computer. This 
-    makes it difficult to thoroughly investigate this instance.
+    The size of our 3 instances vary considerably. The total number of different possible
+    schedules in any problem instance is $|C|^{|R|dt}$ which is far too large to conduct
+    a brute force search. The smallest could potentially be solved exactly, with methods
+    not under consideration here, while the others have search spaces much too large 
+    for any hope of finding a perfect solution. For the largest, which simulates a UofT 
+    exam schedule, a single evaluation of the fitness function takes 0.2 seconds on on 
+    a typical desktop computer. This makes it difficult to thoroughly investigate this 
+    instance.
 
-    The large data set was created with parameters corresponding to the fall 2009 
+    The large problem instance was created with parameters corresponding to the fall 2009 
 	University of Toronto exam schedule, which is the size we expect for a large 
@@
 	\end{small}
-	\caption{The six data sets used in our project. $\#$Maj is the number of different 
+	\caption{The six problem instances used in our project. $\#$Maj is the number of different 
 	         majors used and \textsc{SSS} is an approximation of the search space size.}
@@
   	  \includegraphics[width=8cm]{SMALL-TYPE2_FullAverageStats}
-      \caption{Comparing the four searches on a small data set}
+      \caption{Search methods on small problem instance}
 	\end{figure}
@@
   	  \includegraphics[width=8cm]{HUGE-TYPE2_FullAverageStats}
-      \caption{Comparing the four searches on a large data set}
+      \caption{Search methods on large problem instance}
 	\end{figure}
